### 2025-01-22 main.tex
initial commit

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[hidelinks]{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsmath,amssymb,amsthm}
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{mathrsfs}
\usepackage{graphicx}
\usepackage{doi}
\usepackage{acronym}
\usepackage{listings}
\usepackage{tikz}

\usetikzlibrary{trees}

\newacro{abm}[ABM]{Agent-Based Model}
\newacro{cabm}[CABM]{Cellular Agent-Based Model}
\newacro{ca}[CA]{Cellular Automaton}
\newacroplural{ca}[CA]{Cellular Automata}

\title{Rust for High-Performance Computing}

%\date{September 9, 1985}	% Here you can change the date presented in the paper title
%\date{} 					% Or removing it

\author{
    \href{https://orcid.org/0009-0001-0613-7978}{
        \includegraphics[scale=0.06]{orcid.pdf}
        \hspace{1mm}Jonas Pleyer
    }
    \thanks{
        \href{https://jonas.pleyer.org}{jonas.pleyer.org},
        \href{https://cellular-raza.com}{cellular-raza.com}
    }\\
	Freiburg Center for Data-Analysis and Modeling\\
	University of Freiburg\\
	\texttt{jonas.pleyer@fdm.uni-freiburg.de} \\
	%% examples of more authors
	\And
	\href{https://orcid.org/0000-0002-6371-4495}{
        \includegraphics[scale=0.06]{orcid.pdf}
        \hspace{1mm}Christian Fleck
    }\\
	Freiburg Center for Data-Analysis and Modeling\\
	University of Freiburg
}

% Uncomment to remove the date
%\date{}

% Uncomment to override  the `A preprint' in the header
\renewcommand{\headeright}{Preprint}
%\renewcommand{\undertitle}{Technical Report}
\renewcommand{\shorttitle}{Rust for High-Performance Computing}

\usepackage{enumitem}
\setlist{nolistsep}

%%% Add PDF metadata to help others organize their library
%%% Once the PDF is generated, you can check the metadata with
%%% $ pdfinfo template.pdf
\hypersetup{
pdftitle={Rust for High-Performance Computing},
pdfsubject={q-bio.NC, q-bio.QM},
pdfauthor={Jonas Pleyer, Christian Fleck},
pdfkeywords={},
}

% Change numbering of equations
% \numberwithin{equation}{section}

% MAKE TITLES IN THEOREMS BOLD
\makeatletter
\def\th@plain{%
  \thm@notefont{}% same as heading font
  \itshape % body font
}
\def\th@definition{%
  \thm@notefont{}% same as heading font
  \normalfont % body font
}
\makeatother

\begin{document}
\maketitle

%###################################################################################################
\begin{abstract}
\end{abstract}

% keywords can be removed
\keywords{Rust \and HPC \and High-Performance Computing \and Programming}

%###################################################################################################
\section{Introduction}
\label{section:introduction}

%###################################################################################################
\section{Ergonomics }
\label{section:debugging}

%###################################################################################################
\section{Ecosystem and Libraries}

%---------------------------------------------------------------------------------------------------
\subsection{Numerics}
\label{subsection:numerics}

\begin{enumerate}
    \item ODE solvers~\cite{Renevey2024}
    \item PDE solvers
    \item Fluid Dynamics: Salva (2D \& 3D)~\cite{Crozet2024}
    \item Physics Engine: Rapier (2D \& 3D)~\cite{Crozet2025}
\end{enumerate}

%---------------------------------------------------------------------------------------------------
\subsection{Storage}
\label{subsection:storage}

\begin{enumerate}
    \item serde ecosystem~\cite{TolnaySerde2024,TolnaySerdeJson2025}
\end{enumerate}

%---------------------------------------------------------------------------------------------------
\subsection{Debugging}
\label{subsection:debugging}

%---------------------------------------------------------------------------------------------------
\subsection{GPGPU Programming}
\label{subsection:gpgpu}

\begin{enumerate}
    \item wgpu~\cite{Fitzgerald2025}
\end{enumerate}

%###################################################################################################
\section{Examples}
\label{section:examples}

\begin{enumerate}
    \item Physics~\cite{}
    \item Biology~\cite{Pleyer2024}
    \item Bioinformatics~\cite{}
\end{enumerate}

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}

### 2025-08-20 crate-info-puller/src/main.rs
add crate to pull information from crates.io

@@ -1,0 +1,61 @@
+use anyhow::Context;
+
+#[async_std::main]
+async fn main() -> anyhow::Result<()> {
+    use crates_io_api_wasm_patch::AsyncClient;
+    use reqwest::header::*;
+
+    let crates = ["nalgebra", "ndarray", "cellular_raza", "approx-derive"];
+
+    let mut headers = HeaderMap::new();
+    headers.insert(USER_AGENT, HeaderValue::from_str("gobbler")?);
+
+    let client1 = reqwest::Client::builder()
+        .default_headers(headers)
+        .build()?;
+    let client =
+        AsyncClient::with_http_client(client1.clone(), web_time::Duration::from_millis(1000));
+
+    println!("\\begin{{tabular}}{{l r r}}");
+    println!("    Crate Name       &Weekly Downloads   &Last Update    &Latest Version\\\\");
+    for (n, crate_name) in crates.iter().enumerate() {
+        let cr = client.get_crate(crate_name).await?;
+        let downloads = client.crate_downloads(crate_name).await?;
+        let mut d = downloads.version_downloads;
+        d.sort_by_key(|x| x.date);
+
+        let last_update = cr.crate_data.updated_at;
+        let latest_version = cr
+            .crate_data
+            .versions
+            .and_then(|v| v.into_iter().max())
+            .context("no version provided by crates.io")?;
+
+        let weekly_downloads = {
+            let mut n = 0;
+            let w = d
+                .into_iter()
+                .filter(|x| x.version == latest_version)
+                .map(|di| {
+                    n += 1;
+                    di.downloads
+                })
+                .sum::<u64>();
+            w as f64 / n as f64 * 7.
+        };
+        let latest_version = cr.versions[0].num.clone();
+
+        print!(
+            "    {crate_name:16} &{weekly_downloads:<8.0} &{:10} &{latest_version:10}",
+            last_update.format("%d/%m/%Y")
+        );
+        if n + 1 < crates.len() {
+            println!("\\\\");
+        } else {
+            println!();
+        }
+    }
+    println!("\\end{{tabular}}");
+
+    Ok(())
+}

### 2025-08-20 main.tex
add more ideas
define preliminary structure of paper

@@
 \usepackage{tikz}
+\usepackage[backend=biber,style=ieee]{biblatex}
+
+\addbibresource{references.bib}
 
@@
 
-\title{Rust for High-Performance Computing}
+\title{Rust for Scientific Computing}
 
@@
 
+\pagebreak
+\tableofcontents
+\pagebreak
+
 % keywords can be removed
@@
 %###################################################################################################
-\section{Ergonomics }
-\label{section:debugging}
-
-%###################################################################################################
 \section{Ecosystem and Libraries}
 
-%---------------------------------------------------------------------------------------------------
-\subsection{Numerics}
-\label{subsection:numerics}
-
 \begin{enumerate}
-    \item ODE solvers~\cite{Renevey2024}
+    \item Linear Algebra \& Arrays (nalgebra, ndarray, ...)
+    \item Statistics \& Data Analysis (statrs, ndarray-stats, smartcore, ...)
+    \item Symbolic \& other Math (num, rug, speciyl, symcc)
+    \item ODE solvers~\cite{Renevey2024}, Diffsol
     \item PDE solvers
@@
     \item Physics Engine: Rapier (2D \& 3D)~\cite{Crozet2025}
+    \item Meshing (honeycomb)
+    \item Plotting (plotters, plotlib, gnuplot bindings, vtk-rs)
+    \item Utilities (rayon, lapack)
+    \item GPU (rust-cuda, wgpu)
+    \item Datastructures
+    \item Optimization
 \end{enumerate}
 
-%---------------------------------------------------------------------------------------------------
-\subsection{Storage}
-\label{subsection:storage}
+\subsection{Foundational Components}
+\subsubsection{Linear Algebra \& Arrays}
+(nalgebra, ndarray) (dense, sparse matrices, solvers)
 
-\begin{enumerate}
-    \item serde ecosystem~\cite{TolnaySerde2024,TolnaySerdeJson2025}
-\end{enumerate}
+\subsubsection{Numerical Methods}
+(floating point arithmetic, etc.)
 
-%---------------------------------------------------------------------------------------------------
-\subsection{Debugging}
-\label{subsection:debugging}
+\subsubsection{Statistics \& Probability}
+(statsrs, ndarray-stats, smartcore)
 
-%---------------------------------------------------------------------------------------------------
-\subsection{GPGPU Programming}
-\label{subsection:gpgpu}
+\subsubsection{Symbolic \& Algebraic Computing}
+(num, rug, speciyl, symcc, enzyme)
 
-\begin{enumerate}
-    \item wgpu~\cite{Fitzgerald2025}
-\end{enumerate}
+\subsubsection{Datastructures}
+(graphs, etc.)
+
+\subsection{Computational Methods}
+\subsubsection{Optimization}
+\subsubsection{Differential Equations}
+(ODEs, PDEs)
+
+\subsubsection{Physics Engines}
+(Fluid Dynamics, Manybody)
+
+\subsubsection{Numerical Integration \& Differentiation}
+(quadrature, monte-carlo, symbolic diff, enzyme)
+
+\subsubsection{(Probabilistic Algorithms)}
+
+\subsection{High-Performance \& Scalable Computing}
+\subsubsection{Parallel Computing}
+(rayon)
+
+\subsubsection{GPU \& Accelerators}
+(rust-cuda, wgpu~\cite{Fitzgerald2025})
+
+\subsubsection{Sparse \& Structured Computations}
+(large databases)
+
+\subsubsection{Debugging}
+(Rust Errors as values, tracing, logging) ecosystem
+
+\subsubsection{Storage}
+(serde, json, xml, ron, ...)
+
+\subsection{Visualization}
+\subsubsection{Plotting (2D)}
+\subsubsection{3D Rendering}
+\subsubsection{Interactive}
+(dioxus)
 
 %###################################################################################################
-\section{Examples}
-\label{section:examples}
+\section{Applications}
+\label{section:applications}
 
-\begin{enumerate}
-    \item Physics~\cite{}
-    \item Biology~\cite{Pleyer2024}
-    \item Bioinformatics~\cite{}
-\end{enumerate}
+\subsection{Physics \& Engineering}
+\subsection{Biology \& Chemistry}
+\cite{Pleyer2024}, Computational Oncology~\cite{Köster2025}, Proteomics~\cite{Anechitoaie2024}
+\subsection{Climate \& Earth Sciences}
+\subsection{Economics \& Finance}
+\subsection{Social Sciences}
 
-\bibliographystyle{IEEEtran}
-\bibliography{references}
+%###################################################################################################
+\section{Discussion}
+\label{section:discussion}
+
+%###################################################################################################
+\section{Conclusion}
+\label{section:conclusion}
+
+\printbibliography
 

### 2025-08-22 joss-info-puller/src/main.rs
add issue puller for JOSS Journal

@@ -1,0 +1,59 @@
+#[tokio::main]
+async fn main() {
+    let octo = octocrab::instance();
+
+    let labels = ["published".to_string()];
+
+    let mut page = octo
+        .issues("openjournals", "joss-reviews")
+        .list()
+        .labels(&labels)
+        .state(octocrab::params::State::Closed)
+        .per_page(100)
+        .send()
+        .await
+        .unwrap();
+
+    let mut k = 0;
+    loop {
+        for issue in &page {
+            let mut subpage = octo
+                .issues("openjournals", "joss-reviews")
+                .list_comments(issue.number)
+                .per_page(100)
+                .send()
+                .await
+                .unwrap();
+            'subpage_loop: loop {
+                for comment in &subpage {
+                    if comment
+                        .body
+                        .as_ref()
+                        .is_some_and(|x| x.contains("Rust") || x.contains("rust"))
+                    {
+                        println!("{k:4} {}", issue.title);
+                        k += 1;
+                        break 'subpage_loop;
+                    }
+                }
+
+                subpage = match octo
+                    .get_page::<octocrab::models::issues::Comment>(&subpage.next)
+                    .await
+                    .unwrap()
+                {
+                    Some(next_subpage) => next_subpage,
+                    None => break,
+                }
+            }
+        }
+        page = match octo
+            .get_page::<octocrab::models::issues::Issue>(&page.next)
+            .await
+            .unwrap()
+        {
+            Some(next_page) => next_page,
+            None => break,
+        }
+    }
+}

### 2025-08-22 joss-info-puller/src/main.rs
also search for Rust label

@@ -2,7 +2,7 @@
 async fn main() {
     let octo = octocrab::instance();
 
-    let labels = ["published".to_string()];
+    let labels = ["Rust".to_string(), "published".to_string()];
 
     let mut page = octo
         .issues("openjournals", "joss-reviews")

### 2025-08-22 figures/pypi-org-used-languages.py
add figures for paper

@@ -1,0 +1,53 @@
+import matplotlib.pyplot as plt
+
+
+def prepare_fig_ax():
+    plt.rcParams.update(
+        {
+            "font.family": "Courier New",  # monospace font
+            "font.size": 20,
+            "axes.titlesize": 20,
+            "axes.labelsize": 20,
+            "xtick.labelsize": 20,
+            "ytick.labelsize": 20,
+            "legend.fontsize": 20,
+            "figure.titlesize": 20,
+        }
+    )
+    fig, ax = plt.subplots(figsize=(8, 6))
+    fig.subplots_adjust(
+        top=0.9,
+        left=0.2,
+        right=0.8,
+        bottom=0.15,
+    )
+    ax.grid(True, which="major", linestyle="-", linewidth=0.75, alpha=0.25)
+    ax.minorticks_off()
+    ax.grid(True, which="minor", linestyle="-", linewidth=0.25, alpha=0.15)
+    ax.set_axisbelow(True)
+    return fig, ax
+
+
+data = [
+    ("Go", 20 * 3, "#00acd7"),
+    ("Java", 20 * 10, "#c12f30"),
+    ("Fortran", 20 * 13, "#744e97"),
+    ("C", 20 * 101, "#659bd3"),
+    ("C++", 20 * 119, "#00599d"),
+    ("Rust", 20 * 158, "#CE412B"),
+]
+
+fig, ax = prepare_fig_ax()
+
+ax.barh(
+    [d[0] for d in data],
+    [d[1] for d in data],
+    color=[d[2] for d in data],
+    align="center",
+)
+ax.set_xlabel("(B) Number of Packages")
+ax.set_title("Packages at pypi.org")
+
+fig.savefig("figures/pypi-org-used-languages.png")
+fig.savefig("figures/pypi-org-used-languages.pdf")
+fig.savefig("figures/pypi-org-used-languages.svg")

### 2025-08-22 figures/stackoverflow-popular-languages.py

@@ -1,0 +1,62 @@
+import matplotlib as mpl
+import importlib
+
+pypi_mod = importlib.import_module("pypi-org-used-languages", package=None)
+
+fig, ax = pypi_mod.prepare_fig_ax()
+
+groups = ["Learners", "Professionals", "Total"]
+data_individual = [
+    ("Fortran", [2, 0.9, 1.4], "#744e97"),
+    ("Rust", [23.1, 14.5, 14.8], "#CE412B"),
+    ("Go", [13.1, 17.4, 16.4], "#00acd7"),
+    ("C", [48, 19.1, 22], "#659bd3"),
+    ("C++", [44.6, 21.8, 23.5], "#00599d"),
+    ("Java", [40.8, 29.6, 29.4], "#c12f30"),
+]
+
+data = []
+for d in data_individual:
+    for n, group in enumerate(groups):
+        color = mpl.colors.to_rgba(d[2], alpha=(n + 1) / len(groups))
+        data.append((f"{d[0]} ({group})", d[1][n], color))
+
+labels = [
+    "" if i % 3 != 1 else data_individual[int((i - 1) / 3)][0]
+    for i, d in enumerate(data)
+]
+ax.barh(
+    [d[0] for d in data],
+    [d[1] for d in data],
+    color=[mpl.colors.to_rgba(d[2]) for d in data],
+    label=labels,
+)
+ax.set_yticklabels(labels)  # , horizontalalignment="left")
+
+# handles, labels = ax.get_legend_handles_labels()
+# print(handles[0])
+handles = [
+    mpl.patches.Rectangle(
+        xy=(0, 0.6),
+        width=0.9,
+        height=0.8,
+        angle=0,
+        facecolor=mpl.colors.to_rgba("#000", alpha=(len(groups) - i) / len(groups)),
+    )
+    for i in range(len(groups))
+]
+ax.legend(
+    handles,
+    groups[::-1],
+    ncol=3,
+    frameon=False,
+    loc="upper center",
+    bbox_to_anchor=(0.5, 1.15),
+)
+
+# ax.set_title("Language Popularity (stackoverflow.co)")
+ax.set_xlabel("(A) Language Usage [%]")
+
+fig.savefig("figures/stackoverflow-popular-languages.png")
+fig.savefig("figures/stackoverflow-popular-languages.pdf")
+fig.savefig("figures/stackoverflow-popular-languages.svg")

### 2025-08-22 main.tex
include figures in text
add first list of crates
- still incomplete
- probably redundant
- contains some wrongly formatted results

@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{figures/stackoverflow-popular-languages.pdf}%
+    \includegraphics[width=0.5\textwidth]{figures/pypi-org-used-languages.pdf}
+    \caption{TODO}
+\end{figure}
+
 %###################################################################################################
@@
 \subsubsection{Parallel Computing}
-(rayon)
+(rayon, compare to openmp, MPI bindings)
 
 \subsubsection{GPU \& Accelerators}
-(rust-cuda, wgpu~\cite{Fitzgerald2025})
+(rust-cuda, cudarc, wgpu~\cite{Fitzgerald2025}, opencl, mention SPIRV)
 
@@
 
+\subsection{Creature Comforts}
 \subsubsection{Debugging}
@@
 (serde, json, xml, ron, ...)
+
+\subsubsection{Bindings}
+(pyo3, maturin, C, cxxbridge, cmake)
 
@@
 
+%###################################################################################################
+\pagebreak
+\newcounter{supplementSection}
+\renewcommand{\thesubsection}{S\arabic{supplementSection}}
+\newcommand{\supplement}[1]{%
+    \stepcounter{supplementSection}%
+    \subsection{#1}
+}
+\renewcommand{\thesection}{}
+
+\section{Supplementary Material}
+\supplement{List of all crates}
+
+\paragraph{Scientific Computing in Rust monthly Newsletter}
+
+approx-derive\\
+A procedural macro that automatically implements approximate equality traits for
+floating-point types and custom structs. It builds on the approx crate to make testing numerical
+code easier. This helps ensure tolerance-based equality checks are consistent and less error-prone.
+
+approxim\\
+A library for approximation algorithms in numerical computing. It provides tools for
+approximating functions, datasets, or computations efficiently. The focus is on balancing precision
+with performance in scientific workflows.
+
+bacon-sci\\
+A scientific computing library built on Rust that provides a foundation for numerical
+and algebraic operations. It aims to be a lightweight alternative to larger ecosystems. Its design
+emphasizes safety, performance, and modularity.
+
+cargo-upgrades\\
+A Cargo subcommand for upgrading dependencies in a Rust project. It automates
+version checking and updates Cargo.toml accordingly. This reduces the manual overhead of keeping
+scientific projects up to date.
+
+EGObox\\
+A library for evolutionary game theory and agent-based modeling. It allows researchers to
+explore strategies, equilibria, and dynamics in multi-agent systems. Designed with flexibility to
+model diverse game-theoretic scenarios.
+
+faer\\
+A linear algebra library for Rust with focus on high-performance matrix operations. It
+provides dense and sparse routines with attention to numerical stability. The crate aims to become
+a foundation for scientific and machine-learning applications in Rust.
+
+fact.rs\\
+A crate for fast matrix factorizations and related linear algebra decompositions. It
+provides algorithms like LU, QR, and eigenvalue methods. Its goal is to complement crates like faer
+with specialized routines.
+
+globalsearch\\
+A Rust crate for global optimization and search algorithms. It supports solving
+complex optimization problems where local minima are common. Useful in scientific simulations and
+computational experiments.
+
+indexing\_fmt\\
+A utility crate for formatting indices and ranges in scientific computing contexts.
+It makes working with slices and multi-dimensional arrays more ergonomic. This helps in debugging
+and presenting results.
+
+integraal\\
+A symbolic and numerical integration library for Rust. It supports computing definite
+and indefinite integrals for a variety of functions. Its goal is to provide a reliable integration
+toolkit for researchers.
+
+kifmm-rs\\
+A Rust implementation of the kernel-independent fast multipole method (FMM). It
+accelerates particle simulations and large-scale interactions in physics. The focus is on
+high-performance computing and scalability.
+
+ndgrid\\
+A crate for generating multi-dimensional grids, similar to MATLAB’s ndgrid. It is
+especially useful in scientific computing for parameter sweeps and simulations. Provides ergonomic
+grid construction for Rust’s array ecosystem.
+
+ndelement\\
+A crate for manipulating individual elements in multi-dimensional arrays. It offers
+convenient indexing and iteration across dimensions. Designed to complement libraries like ndarray.
+
+ndarray\\
+A core Rust crate for n-dimensional arrays, inspired by NumPy. It provides efficient
+array structures, broadcasting, slicing, and linear algebra support. Widely used as a foundation
+for many Rust scientific computing libraries.
+
+ploc\\
+A parallelized algorithm library for localization and clustering tasks. Focuses on scalable
+implementations of clustering approaches. Intended for large datasets in scientific applications.
+
+Polars\\
+A fast DataFrame library for Rust, with APIs also available in Python. It provides
+powerful data manipulation, joins, and group-by operations with performance close to Apache Arrow.
+Designed for data science, analytics, and scientific workflows.
+
+Quant-Iron\\
+A crate for quantitative finance and scientific modeling. It implements financial
+instruments, stochastic processes, and numerical solvers. Designed to bridge scientific computing
+with financial applications.
+
+paradis\\
+A discrete dislocation dynamics simulator written in Rust. It is used in materials
+science to model dislocation behavior under stress. The library emphasizes performance and
+scalability for large simulations.
+
+repgenerate\\
+A tool for generating reproducible reports from computational experiments. It
+integrates with Rust code to ensure scientific workflows can be repeated. Useful for open and
+transparent research.
+
+rsmpi\\
+Rust bindings to the Message Passing Interface (MPI). It enables parallel computing across
+distributed systems, a key component in HPC. Provides safe abstractions while maintaining low-level
+control.
+
+serde\\
+A widely used Rust serialization and deserialization framework. It supports many formats
+like JSON, YAML, and custom binary encodings. In scientific computing, it helps with data exchange
+and configuration management.
+
+sequenceprofiler\\
+A bioinformatics library for profiling biological sequences. Provides tools for
+analyzing DNA, RNA, or protein sequences. Aimed at high-performance sequence analysis in Rust.
+
+zarrs\\
+A Rust implementation of the Zarr data format for chunked, compressed N-dimensional arrays.
+It allows efficient storage and manipulation of large scientific datasets. Especially useful in
+machine learning and climate science.
+
+\paragraph{JOSS}
+
+fastatomstruct\\
+A Rust library for high-performance analysis of atomic systems. It provides tools
+for structural and dynamical computations, enabling simulations and data analysis in computational
+chemistry and materials science. Optimized for speed, it can handle large datasets efficiently.
+
+grepq\\
+A Rust tool designed to process and filter FASTQ files efficiently. It matches sequences
+against sets of regular expressions, making it useful for high-throughput sequencing data analysis.
+Its focus is on speed and flexibility for genomic workflows.
+
+pengWann\\
+A library that computes chemical bonding descriptors using Wannier functions. It aids in
+understanding electronic structure and bonding characteristics in materials. Suitable for quantum
+chemistry and materials science applications.
+
+wrenfold\\
+A Rust library for symbolic computation and code generation tailored to robotics
+applications. It allows automatic generation of control and kinematics code from symbolic
+expressions. Supports efficient development of robotic algorithms.
+
+kifmm-rs\\
+A Rust framework implementing the kernel-independent fast multipole method (FMM). Used
+to accelerate N-body computations in physics and engineering. Optimized for performance and
+scalable to large simulations.
+
+startinpy\\
+A Python package for modeling and processing 2.5D terrains. Supports triangulated
+meshes and geospatial analysis. Useful in geoscience, terrain analysis, and visualization.
+
+Back to sequences\\
+A tool for tracing the origins of k-mers in sequencing data. Helps analyze and
+understand sequence composition and frequency. Useful for genomics and bioinformatics research.
+
+sourmash v4\\
+A bioinformatics toolkit for genome and metagenome analysis. Provides fast
+comparison, search, and analysis of sequencing data. Supports scalable workflows with large
+datasets.
+
+extendr\\
+A Rust library that facilitates seamless integration with the R programming language.
+Allows calling Rust functions from R and vice versa. Focuses on performance and ease of use for
+scientific computing.
+
+Pywaterflood\\
+A Python library for analyzing well connectivity in hydrogeology or petroleum
+engineering. Uses capacitance-resistance modeling to simulate flow and connectivity. Supports
+data-driven analysis for reservoir studies.
+
+Gibbs Sea Water Oceanographic Toolbox\\
+A Rust implementation of the TEOS-10 toolbox for
+oceanographic computations. Provides tools for thermodynamic and salinity calculations in seawater.
+Useful for oceanographers and climate researchers.
+
+Raphtory\\
+A toolkit for building and analyzing temporal graphs. Supports both Rust and Python
+interfaces. Enables high-performance analytics on time-evolving network data.
+
+HDT\\
+A Rust library for working with HDT, a binary compression format for RDF data. Optimizes
+storage and query of large knowledge graphs. Useful for semantic web and linked data applications.
+
+egobox\\
+A Rust toolbox for performing global optimization. Implements algorithms for efficiently
+exploring high-dimensional spaces. Useful in engineering, machine learning, and simulation-based
+optimization.
+
+Fast k-medoids Clustering\\
+A high-performance implementation of k-medoids clustering in Rust, with
+Python bindings. Supports scalable clustering of large datasets. Ideal for machine learning and
+data analysis tasks.
+
+FEM\_2D\\
+A Rust library for 2D finite element simulations. Supports hp-refinement for adaptive
+mesh accuracy. Designed for engineering, physics, and numerical analysis applications.
+
+retworkx\\
+A Python graph library with Rust backend for performance. Supports graph creation,
+traversal, and analysis. Useful in network science, bioinformatics, and algorithm development.
+
+Rasusa\\
+A bioinformatics tool for subsampling sequencing reads. Allows control of coverage in
+genomic datasets. Helps optimize computational pipelines and reduce data size for analysis.
+
+Sepia\\
+A Rust tool for classifying sequencing reads by taxonomy. Provides high-performance
+analysis for metagenomics studies. Helps researchers quickly categorize large datasets.
+
+RustBCA\\
+A Rust library implementing the binary collision approximation (BCA) for ion-material
+simulations. Enables efficient modeling of particle interactions in materials. Used in physics and
+materials research.
+
+City2BA\\
+A Rust package for generating synthetic bundle adjustment datasets. Useful in computer
+vision and photogrammetry. Helps test and benchmark 3D reconstruction algorithms.
+
+\paragraph{Scientific Computing in Rust 2025 talks}
+
+Diffsol\\
+A crate for solving differential equations.
+
+Pixi\\
+The missing companion to Cargo.
+
+ploc\\
+Efficient point location with ploc.
+
+NPB-Rust\\
+NAS parallel benchmarks in Rust.
+
+cellular\_raza\\
+Cellular agent-based modeling from a clean slate.
+
+ninterp\\
+Numerical interpolation in N-dimensions.
+
+Arrow\\
+Arrow, Rust, and cross-language data science tooling.
+
+Juice\\
+Juice your simulations: what science can learn from game development.
+
+eVaiutilities\\
+Research data management post eVai.
+
+Honeycomb\\
+Combinatorial maps implementation for meshing applications.
+
+Stochastic\\
+Stochastic calculus in Rust.
+
+Exponential\\
+Exponential time integration for stiff systems in Rust.
+
+Rust\\
+Rust for chip design algorithms and EDA software.
+
+Using\\
+Using Rust to mitigate breaking changes in scientific software.
+
+The\\
+The pleasure and pain of developing cross-platform CPU vector code in Rust.
+
+Deimos\\
+Open-source scientific data acquisition \& laboratory controls.
+
+Rust\\
+Rust is RAD and this is why.
+
+Diffsol\\
+A crate for solving differential equations.
+
+Pixi\\
+The missing companion to Cargo.
+
+ploc\\
+Efficient point location with ploc.
+
+\paragraph{Scientific Computing in Rust 2024 talks}
+
+AcoDyn\\
+Fluid dynamics on the GPU powered by Rust.
+
+RLST\\
+Linear algebra frameworks in Rust for the solution of partial differential equations.
+
+Raphtory\\
+Temporal graphs from Scala to Rust.
+
+Bless\\
+Transparently logging program outputs.
+
+sophus-rs\\
+A fast electromagnetics library with Rust and Python.
+
+sophus-rs\\
+A fast electromagnetics library with Rust and Python.
+
+Rootfinders\\
+Rootfinders for Rust.
+
+Rstats\\
+Multidimensional data analysis in Rust.
+
+extendr\\
+Frictionless bindings for R and Rust.
+
+Rusting\\
+Rusting RSRS.
+
+Physics\\
+Physics simulations in Bevy.
+
+An\\
+An embedded memory cache for persistent memory access.
+
+Polars\\
+Polars in the factory: when databases trump collections for modeling.
+
+Perpetaul\\
+Perpetaul: a hyperparameter-free gradient boosting machine.
+
+Writing\\
+Writing maths software in Rust after 20 years of C++: graph isomorphism and groups.
+
+Productive\\
+Productive and reliable constitutive modeling in Rust.
+
+Efficient\\
+Efficient 1-bit matrix approximations.
+
+Developing\\
+Developing Rust-based R packages using the Roxido framework.
+
+Rusph\\
+Rusph: a SPH astrophysical simulation code in the Rust programming language.
+
+Back\\
+Back to basics: rebuilding computational biomechanics crate by crate.
+
+Cargo\\
+Cargo +GPU build: An early outlook.
+
+Writing\\
+Writing a grid library for finite and boundary element methods.
+
+\paragraph{Scientific Computing in Rust 2023 talks}
+
+jiro\_nn\\
+Low-friction high-detail Deep Learning framework in Rust
+
+\url{https://chrisproject.org/}\\
+ChRIS is an open-source platform for computational research and medicine.
+
+struqture\\
+Represent quantum mechanical operators, Hamiltonians and open quantum systems.
+\url{https://quantumsimulations.de/}
+
+salso, caviarpd, \& fangs\\
+R extensions written in Rust
+
+faer-rs\\
+A linear algebra foundation for the Rust programming language.
+
+Lamellar\\
+A Rust-based asynchronous tasking and PGAS runtime for high-performance computing.
+
+Sage\\
+Enables fast proteomics searching and quantification at scale.
+
+Bernstein–Bézier\\
+Finite elements for RMM in Rust. ResearchGate
+
+Roqoqo\\
+A quantum computing toolkit in Rust.
+
+Lace\\
+Bayesian tabular data analysis for Rust (and Python).
+
+Rudolph\\
+Improving productivity of medical research with Rust CLI tools.
+
+Rstats\\
+Multidimensional data analysis in Rust.
+
+extendr\\
+Frictionless bindings for R and Rust.
+
+Rusting\\
+Rusting RSRS.
+
+Physics\\
+Physics simulations in Bevy.
+
+An\\
+An embedded memory cache for persistent memory access.
+
+Polars\\
+Polars in the factory: when databases trump collections for modeling.
+
+Perpetaul\\
+Perpetaul: a hyperparameter-free gradient boosting machine.
+
+Writing\\
+Writing maths software in Rust after 20 years of C++: graph isomorphism and groups.
+
+Productive\\
+Productive and reliable constitutive modeling in Rust.
+
+Efficient\\
+Efficient 1-bit matrix approximations.
+
+Developing\\
+Developing Rust-based R packages using the Roxido framework.
+
+Rusph\\
+Rusph: a SPH astrophysical simulation code in the Rust programming language.
+
+Back\\
+Back to basics: rebuilding computational biomechanics crate by crate.
+
+Cargo\ \ Cargo +GPU build: An early outlook.
+
+Writing\\
+Writing a grid library for finite and boundary element methods.
+
+\paragraph{FAANG Companies}
+
+Amazon (AWS)\\
+AWS developed Firecracker, a microVM monitoring service powering AWS Lambda and AWS Fargate.
+Firecracker is built primarily in Rust to offer fast cold boot times and secure workload isolation.
+
+Rust-supported runtimes like AWS Lambda also allow developers to write serverless functions using
+Rust. DEV Community Litslink
+
+Google\\
+Fuchsia, Google's modern operating system, uses Rust extensively for system components such as
+drivers and services to enhance security and reliability.
+
+Android integration: Rust has been adopted within the Android Open Source Project (AOSP) to improve
+memory safety in critical system components. DEV Community Litslink
+
+Meta (Facebook)\\
+Mononoke, Facebook’s large-scale source control backend for Mercurial, was written in Rust to
+improve performance and reliability in its monorepo.
+
+Facebook employs Rust in developing CLI tools, distributed systems, and cron jobs. Engineers write
+millions of lines of Rust internally. Engineering at Meta Facebook Developers
+
+Apple\\
+Apple’s Cloud Traffic team and backend infrastructure have been shifting from C to Rust, with new
+functionality being primarily built in Rust. This change was driven by performance, security, and
+developer efficiency goals. Reddit
+
+Microsoft\\
+Although not part of FAANG, it’s noteworthy—Microsoft is rewriting parts of the Windows kernel and
+system components in Rust (e.g., DWriteCore, GDI Regions), prioritizing memory safety.
+
+Azure services—like Azure IoT Edge—also incorporate Rust. Litslink Wikipedia
+
 \end{document}

### 2025-08-22 figures/stackoverflow-desired-language.py
add more plots

@@ -1,0 +1,67 @@
+import matplotlib as mpl
+import importlib
+
+pypi_mod = importlib.import_module("pypi-org-used-languages", package=None)
+
+fig1, ax1 = pypi_mod.prepare_fig_ax()
+fig2, ax2 = pypi_mod.prepare_fig_ax()
+
+years = [2017, 2018, 2019, 2020, 2021, 2022, 2023, 2024, 2025]
+
+# Name, Loved, Wanted, Color
+languages = [
+    (
+        "Rust",
+        [73.1, 78.9, 83.5, 86.1, 86.98, 86.73, 84.66, 82.2, 72.4],
+        [6.6, 8.3, 9.5, 14.6, 14.09, 17.6, 30.56, 28.7, 29.2],
+        "#CE412B",
+    ),
+    (
+        "Go",
+        [63.3, 65.6, 67.9, 62.3, 62.74, 64.58, 62.45, 67.7, 56.5],
+        [13.5, 16.2, 15.0, 17.9, 14.54, 16.41, 20.59, 23.1, 23.4],
+        "#00acd7",
+    ),
+    (
+        "C",
+        [41.7, None, 42.5, 33.1, 39.56, 39.68, 43.29, 47.4, 45.0],
+        [6.4, 5.9, 5.0, 4.3, 4.52, 4.34, 11.51, 13.9, 14.5],
+        "#659bd3",
+    ),
+    (
+        "C++",
+        [52.0, 46.7, 52.0, 43.4, 49.24, 48.39, 49.77, 53.1, 46.6],
+        [11.8, 10.2, 9.1, 8.6, 8.8, 7.67, 16.35, 18.3, 16.7],
+        "#00599d",
+    ),
+    (
+        "Java",
+        [50.5, 50.7, 53.4, 44.1, 47.15, 45.75, 44.11, 47.6, 41.8],
+        [11.7, 10.5, 8.3, 8.8, 6.79, 5.6, 16.53, 17.9, 15.8],
+        "#c12f30",
+    ),
+]
+
+for name, loved, wanted, color in languages:
+    ax1.plot(years, loved, label=name, color=color)
+    ax2.plot(years, wanted, label=name, color=color)
+
+labels = ["Most Loved Language", "Most Desired Language"]
+for n, ax in enumerate([ax1, ax2]):
+    ax.set_title(labels[n])
+    ax.set_xlabel("Year")
+    ax.set_ylabel("Developers [%]")
+    ax.legend(
[…]

### 2025-08-22 main.tex
add more notes for introduction

@@
 \label{section:introduction}
+
+General Remarks about languages
+\begin{enumerate}
+    \item Programming languages allow us to express machine instructions in human-readable code.
+    \item Depend on use-case
+    \item Translate human-interpretable code into machine instructions
+    \item tradeoff between high-level view (abstraction) and performance (low level)
+    \item C++ has "zero-cost" abstractions
+\end{enumerate}
+
+Importance of programming in science
+\begin{enumerate}
+    \item New fields of study (computational physics, chemistry, biology, sociology, etc.)
+    \item Cite some groundbreaking examples which were possible due to programming in science (maybe
+          weather)
+    \item
+\end{enumerate}
+
+Judging C/C++
+\begin{enumerate}
+    \item Previous Decades: C/C++ in science
+    \item Limitations of design choices of language become evident
+    \item fragmented ecosystem; good for competition; bad for developer experience
+    \item modern features missing: package manager with package registries, build system
+    \item technical details; mutable by default, templates vs generics, backwards compatiblity
+          becomes a problem ("there is a much simpler langauge within C++ that tries to get out"),
+          macros, no clear ownership rules on a language level, dangling pointers
+    \item still many good principles (such as RAII) and design principles/concepts (cite gang of
+          four)
+\end{enumerate}
+
+The need for modern languages
+\begin{enumerate}
+    \item Java aimed to replace C due to pointers
+    \item Go also aims to replace it, go is garbage-collected; thus cannot be as well-performing as
+          C++ or C
+    \item Python, Javascript for high-level abstractions ("as glue"); no performance howerver
+\end{enumerate}
+
+Elaborate on why Rust solves many problems
+\begin{enumerate}
+    \item safety guarantees; ensures that code which we write actually performs what we set out to
+          do
+    \item no dangling pointers
+    \item zero-cost abstrations
+    \item compiled without garbage-collector
+    \item tooling: modern ecosystem; package manager, build system, package registry
+    \item strong type system which is generally considered sound
+    \item macro system gets type-checked
+    \item very good C-interop; utilize existing software
+\end{enumerate}
+
+Show that many companies are starting to adopt Rust
+\begin{enumerate}
+    \item Amazon (AWS)\\
+          AWS developed Firecracker, a microVM monitoring service powering AWS Lambda and AWS
+          Fargate.
+          Firecracker is built primarily in Rust to offer fast cold boot times and secure
+          workload isolation.
+    \item Rust-supported runtimes like AWS Lambda also allow developers to write serverless
+          functions using Rust.
+    \item Google\\
+          Fuchsia, Google's modern operating system, uses Rust extensively for system components
+          such as drivers and services to enhance security and reliability.
+    \item Google (Android)\\
+          Android integration: Rust has been adopted within the Android Open Source Project (AOSP)
+          to improve memory safety in critical system components. DEV Community Litslink
+    \item Meta (Facebook)\\
+          Mononoke, Facebook’s large-scale source control backend for Mercurial, was written in
+          Rust to improve performance and reliability in its monorepo.
+    \item Meta (Facebook)\\
+          Facebook employs Rust in developing CLI tools, distributed systems, and cron jobs.
+          Engineers write millions of lines of Rust internally.
+    \item Apple\\
+          Apple’s Cloud Traffic team and backend infrastructure have been shifting from C to Rust,
+          with new functionality being primarily built in Rust. This change was driven by
+          performance, security, and developer efficiency goals. Reddit
+    \item Microsoft\\
+          Although not part of FAANG, it’s noteworthy—Microsoft is rewriting parts of the Windows
+          kernel and system components in Rust (e.g., DWriteCore, GDI Regions), prioritizing memory
+          safety.
+    \item Azure services-like Azure IoT Edge—also incorporate Rust.
+    \item Linux Kernel\\
+          Only other supported language except from C
+\end{enumerate}
 

### 2025-08-22 main.tex
add new images
add new crates; distribute remaining points onto sections

@@
     \includegraphics[width=0.5\textwidth]{figures/pypi-org-used-languages.pdf}
+    \includegraphics[width=0.5\textwidth]{figures/stackoverflow-loved-language.pdf}%
+    \includegraphics[width=0.5\textwidth]{figures/stackoverflow-desired-language.pdf}
     \caption{TODO}
@@
 \section{Ecosystem and Libraries}
-
-\begin{enumerate}
-    \item Linear Algebra \& Arrays (nalgebra, ndarray, ...)
-    \item Statistics \& Data Analysis (statrs, ndarray-stats, smartcore, ...)
-    \item Symbolic \& other Math (num, rug, speciyl, symcc)
-    \item ODE solvers~\cite{Renevey2024}, Diffsol
-    \item PDE solvers
-    \item Fluid Dynamics: Salva (2D \& 3D)~\cite{Crozet2024}
-    \item Physics Engine: Rapier (2D \& 3D)~\cite{Crozet2025}
-    \item Meshing (honeycomb)
-    \item Plotting (plotters, plotlib, gnuplot bindings, vtk-rs)
-    \item Utilities (rayon, lapack)
-    \item GPU (rust-cuda, wgpu)
-    \item Datastructures
-    \item Optimization
-\end{enumerate}
 
@@
 \subsubsection{Statistics \& Probability}
-(statsrs, ndarray-stats, smartcore)
+(statrs, ndarray-stats, smartcore)
 
@@
 
+ODE solvers~\cite{Renevey2024}, Diffsol
+
 \subsubsection{Physics Engines}
 (Fluid Dynamics, Manybody)
+
+salva (2D \& 3D)~\cite{Crozet2024}
+
+Rapier (2D \& 3D)~\cite{Crozet2025}
 
@@
 \subsubsection{Plotting (2D)}
+plotters, plotlib, gnuplot bindings
+
 \subsubsection{3D Rendering}
+vtk-rs
+
+Meshing (honeycomb)
+
 \subsubsection{Interactive}
@@
 
-RLST\\
-Linear algebra frameworks in Rust for the solution of partial differential equations.
+rlst\\
+The Rust Linear Solver toolbox is an in-development project for dense and sparse linear algebra routines in Rust.
 
@@
 Bernstein–Bézier\\
-Finite elements for RMM in Rust. ResearchGate
+Finite elements for RMM in Rust~\cite{Sky2024}
+
+mwa\_hyperdrive\\
+Calibration software for the Murchison Widefield Array (MWA) radio telescope.
+\url{https://mwatelescope.github.io/mwa_hyperdrive/index.html}
+
+alg\_tools\\
+utility routines and tools for implementing iterative algorithms and (abstract) numerical computing in Rust~\cite{Valkonen2023}.
+\url{https://tuomov.iki.fi/software/alg_tools/}
+
+clarabel\\
+Clarabel is an interior point numerical solver for convex optimization problems using a novel
+homogeneous embedding~\cite{Clarabel_2024}.
+
+fips\\
+\cite{jeggle2023genericframeworkdataracefreemanyparticle}
+\url{https://zenodo.org/records/6757615}
 

### 2025-08-22 main.tex
cleanup

@@
 
-Amazon (AWS)\\
-AWS developed Firecracker, a microVM monitoring service powering AWS Lambda and AWS Fargate.
-Firecracker is built primarily in Rust to offer fast cold boot times and secure workload isolation.
-
-Rust-supported runtimes like AWS Lambda also allow developers to write serverless functions using
-Rust. DEV Community Litslink
-
-Google\\
-Fuchsia, Google's modern operating system, uses Rust extensively for system components such as
-drivers and services to enhance security and reliability.
-
-Android integration: Rust has been adopted within the Android Open Source Project (AOSP) to improve
-memory safety in critical system components. DEV Community Litslink
-
-Meta (Facebook)\\
-Mononoke, Facebook’s large-scale source control backend for Mercurial, was written in Rust to
-improve performance and reliability in its monorepo.
-
-Facebook employs Rust in developing CLI tools, distributed systems, and cron jobs. Engineers write
-millions of lines of Rust internally. Engineering at Meta Facebook Developers
-
-Apple\\
-Apple’s Cloud Traffic team and backend infrastructure have been shifting from C to Rust, with new
-functionality being primarily built in Rust. This change was driven by performance, security, and
-developer efficiency goals. Reddit
-
-Microsoft\\
-Although not part of FAANG, it’s noteworthy—Microsoft is rewriting parts of the Windows kernel and
-system components in Rust (e.g., DWriteCore, GDI Regions), prioritizing memory safety.
-
-Azure services—like Azure IoT Edge—also incorporate Rust. Litslink Wikipedia
-
 \end{document}

### 2025-08-24 crate-info-puller/src/main.rs
restructure program

@@ -1,8 +1,25 @@
-use anyhow::Context;
+use anyhow::{Context, Result};
+use crates_io_api_wasm_patch::{AsyncClient, CrateDownloads, CrateResponse};
 
-#[async_std::main]
-async fn main() -> anyhow::Result<()> {
-    use crates_io_api_wasm_patch::AsyncClient;
+struct CrateData {
+    crate_name: String,
+    crate_response: CrateResponse,
+    downloads: CrateDownloads,
+}
+
+async fn get_data(client: &AsyncClient, crate_name: &str) -> Result<CrateData> {
+    let cre = client.get_crate(crate_name);
+    let dow = client.crate_downloads(crate_name);
+
+    Ok(CrateData {
+        crate_name: crate_name.to_string(),
+        crate_response: cre.await?,
+        downloads: dow.await?,
+    })
+}
+
+#[macro_rules_attribute::apply(smol_macros::main)]
+async fn main() -> Result<()> {
     use reqwest::header::*;
 
     let crates = ["nalgebra", "ndarray", "cellular_raza", "approx-derive"];
@@ -18,14 +35,15 @@ async fn main() -> anyhow::Result<()> {
 
     println!("\\begin{{tabular}}{{l r r}}");
     println!("    Crate Name       &Weekly Downloads   &Last Update    &Latest Version\\\\");
-    for (n, crate_name) in crates.iter().enumerate() {
-        let cr = client.get_crate(crate_name).await?;
-        let downloads = client.crate_downloads(crate_name).await?;
-        let mut d = downloads.version_downloads;
+    for (n, name) in crates.into_iter().enumerate() {
+        let cd = async_compat::Compat::new(get_data(&client, name)).await?;
+
+        let mut d = cd.downloads.version_downloads;
         d.sort_by_key(|x| x.date);
 
-        let last_update = cr.crate_data.updated_at;
-        let latest_version = cr
+        let last_update = cd.crate_response.crate_data.updated_at;
+        let latest_version = cd
+            .crate_response
             .crate_data
             .versions
             .and_then(|v| v.into_iter().max())
@@ -43,10 +61,11 @@ async fn main() -> anyhow::Result<()> {
                 .sum::<u64>();
             w as f64 / n as f64 * 7.
         };
-        let latest_version = cr.versions[0].num.clone();
+        let latest_version = cd.crate_response.versions[0].num.clone();
 
         print!(
[…]

### 2025-08-24 crate-info-puller/src/main.rs
use clap to parse arguments

@@ -18,11 +18,24 @@ async fn get_data(client: &AsyncClient, crate_name: &str) -> Result<CrateData> {
     })
 }
 
+#[derive(clap::Parser, Debug)]
+#[command(version, about, long_about = None)]
+struct Args {
+    /// Comma-separated list of crate names
+    #[arg(short, long, value_delimiter = ',')]
+    crates: Vec<String>,
+    // #[arg(short, long)]
+    // filename: Option<String>,
+}
+
 #[macro_rules_attribute::apply(smol_macros::main)]
 async fn main() -> Result<()> {
+    use clap::Parser;
     use reqwest::header::*;
 
-    let crates = ["nalgebra", "ndarray", "cellular_raza", "approx-derive"];
+    let args = Args::parse();
+    // let crates = ["nalgebra", "ndarray", "cellular_raza", "approx-derive"];
+    let crates = args.crates;
 
     let mut headers = HeaderMap::new();
     headers.insert(USER_AGENT, HeaderValue::from_str("gobbler")?);
@@ -35,7 +48,7 @@ async fn main() -> Result<()> {
 
     println!("\\begin{{tabular}}{{l r r}}");
     println!("    Crate Name       &Weekly Downloads   &Last Update    &Latest Version\\\\");
-    for (n, name) in crates.into_iter().enumerate() {
+    for (n, name) in crates.iter().enumerate() {
         let cd = async_compat::Compat::new(get_data(&client, name)).await?;
 
         let mut d = cd.downloads.version_downloads;

### 2025-08-25 crate-info-puller/src/main.rs
use update crates; implement passing of arguments to CLI

@@ -18,14 +18,21 @@ async fn get_data(client: &AsyncClient, crate_name: &str) -> Result<CrateData> {
     })
 }
 
+/// Query crates.io and generate a summary table
+///
+/// The list of crates is combined in order from
+/// the explicitly given names via the --crates
+/// argument and optionally from the contents of
+/// a file given by the --file argument.
 #[derive(clap::Parser, Debug)]
-#[command(version, about, long_about = None)]
+#[command(version, about, max_term_width = 50)]
 struct Args {
     /// Comma-separated list of crate names
     #[arg(short, long, value_delimiter = ',')]
     crates: Vec<String>,
-    // #[arg(short, long)]
-    // filename: Option<String>,
+    /// Read crate names from lines in a text file
+    #[arg(short, long)]
+    file: Option<String>,
 }
 
 #[macro_rules_attribute::apply(smol_macros::main)]
@@ -34,8 +41,11 @@ async fn main() -> Result<()> {
     use reqwest::header::*;
 
     let args = Args::parse();
-    // let crates = ["nalgebra", "ndarray", "cellular_raza", "approx-derive"];
-    let crates = args.crates;
+    let mut crates = args.crates;
+    if let Some(filename) = args.file {
+        let content = std::fs::read_to_string(filename)?;
+        crates.extend(content.lines().map(String::from));
+    }
 
     let mut headers = HeaderMap::new();
     headers.insert(USER_AGENT, HeaderValue::from_str("gobbler")?);

### 2025-08-25 crate-info-puller/src/main.rs
change description; use generalized formatting routine

@@ -18,48 +18,11 @@ async fn get_data(client: &AsyncClient, crate_name: &str) -> Result<CrateData> {
     })
 }
 
-/// Query crates.io and generate a summary table
-///
-/// The list of crates is combined in order from
-/// the explicitly given names via the --crates
-/// argument and optionally from the contents of
-/// a file given by the --file argument.
-#[derive(clap::Parser, Debug)]
-#[command(version, about, max_term_width = 50)]
-struct Args {
-    /// Comma-separated list of crate names
-    #[arg(short, long, value_delimiter = ',')]
-    crates: Vec<String>,
-    /// Read crate names from lines in a text file
-    #[arg(short, long)]
-    file: Option<String>,
-}
-
-#[macro_rules_attribute::apply(smol_macros::main)]
-async fn main() -> Result<()> {
-    use clap::Parser;
-    use reqwest::header::*;
-
-    let args = Args::parse();
-    let mut crates = args.crates;
-    if let Some(filename) = args.file {
-        let content = std::fs::read_to_string(filename)?;
-        crates.extend(content.lines().map(String::from));
-    }
-
-    let mut headers = HeaderMap::new();
-    headers.insert(USER_AGENT, HeaderValue::from_str("gobbler")?);
-
-    let client1 = reqwest::Client::builder()
-        .default_headers(headers)
-        .build()?;
-    let client =
-        AsyncClient::with_http_client(client1.clone(), web_time::Duration::from_millis(1000));
-
+async fn format_crates<T: AsRef<str>>(client: &AsyncClient, crates: &[T]) -> Result<()> {
     println!("\\begin{{tabular}}{{l r r}}");
     println!("    Crate Name       &Weekly Downloads   &Last Update    &Latest Version\\\\");
     for (n, name) in crates.iter().enumerate() {
-        let cd = async_compat::Compat::new(get_data(&client, name)).await?;
+        let cd = async_compat::Compat::new(get_data(client, name.as_ref())).await?;
 
         let mut d = cd.downloads.version_downloads;
         d.sort_by_key(|x| x.date);
@@ -98,6 +61,46 @@ async fn main() -> Result<()> {
         }
     }
     println!("\\end{{tabular}}");
+    Ok(())
+}
+
+/// Query crates.io and generate a summary table
+///
+/// The crate names are sorted by name before any output is produced.
+#[derive(clap::Parser, Debug)]
+#[command(version, about, max_term_width = 60)]
[…]

### 2025-08-25 crate-info-puller/src/main.rs
print number of dependencies in table

@@ -5,22 +5,26 @@ struct CrateData {
     crate_name: String,
     crate_response: CrateResponse,
     downloads: CrateDownloads,
+    deps: usize,
 }
 
 async fn get_data(client: &AsyncClient, crate_name: &str) -> Result<CrateData> {
-    let cre = client.get_crate(crate_name);
     let dow = client.crate_downloads(crate_name);
+    let crate_response = client.get_crate(crate_name).await?;
+    let num = &crate_response.versions.first().unwrap().num.clone();
+    let deps = client.crate_dependencies(crate_name, num);
 
     Ok(CrateData {
         crate_name: crate_name.to_string(),
-        crate_response: cre.await?,
+        crate_response,
         downloads: dow.await?,
+        deps: deps.await?.len(),
     })
 }
 
 async fn format_crates<T: AsRef<str>>(client: &AsyncClient, crates: &[T]) -> Result<()> {
     println!("\\begin{{tabular}}{{l r r}}");
-    println!("    Crate Name       &Weekly Downloads   &Last Update    &Latest Version\\\\");
+    println!("    Crate Name &Weekly Downloads &Last Update &Latest Version &Dependencies\\\\");
     for (n, name) in crates.iter().enumerate() {
         let cd = async_compat::Compat::new(get_data(client, name.as_ref())).await?;
 
@@ -48,9 +52,10 @@ async fn format_crates<T: AsRef<str>>(client: &AsyncClient, crates: &[T]) -> Res
             w as f64 / n as f64 * 7.
         };
         let latest_version = cd.crate_response.versions[0].num.clone();
+        let n_deps = cd.deps;
 
         print!(
-            "    {:16} &{weekly_downloads:<8.0} &{:10} &{latest_version:10}",
+            "    {:16} &{weekly_downloads:<8.0} &{:10} &{latest_version:10} &{n_deps}",
             cd.crate_name,
             last_update.format("%d/%m/%Y")
         );

### 2025-08-25 crate-info-puller/src/main.rs
use ron storage format to store intermediate results to disk

@@ -1,6 +1,8 @@
 use anyhow::{Context, Result};
 use crates_io_api_wasm_patch::{AsyncClient, CrateDownloads, CrateResponse};
+use serde::{Deserialize, Serialize};
 
+#[derive(Clone, Deserialize, Serialize, Debug)]
 struct CrateData {
     crate_name: String,
     crate_response: CrateResponse,
@@ -9,17 +11,32 @@ struct CrateData {
 }
 
 async fn get_data(client: &AsyncClient, crate_name: &str) -> Result<CrateData> {
-    let dow = client.crate_downloads(crate_name);
-    let crate_response = client.get_crate(crate_name).await?;
-    let num = &crate_response.versions.first().unwrap().num.clone();
-    let deps = client.crate_dependencies(crate_name, num);
+    // First check if data was stored to disk
+    let mut crate_path = std::path::PathBuf::new();
+    crate_path.push("./out/");
+    crate_path.push(crate_name);
 
-    Ok(CrateData {
-        crate_name: crate_name.to_string(),
-        crate_response,
-        downloads: dow.await?,
-        deps: deps.await?.len(),
-    })
+    if crate_path.exists() {
+        let content = std::fs::read_to_string(crate_path)?;
+        let res: CrateData = ron::from_str(&content)?;
+        Ok(res)
+    } else {
+        let dow = client.crate_downloads(crate_name);
+        let crate_response = client.get_crate(crate_name).await?;
+        let num = &crate_response.versions.first().unwrap().num.clone();
+        let deps = client.crate_dependencies(crate_name, num);
+
+        let res = CrateData {
+            crate_name: crate_name.to_string(),
+            crate_response,
+            downloads: dow.await?,
+            deps: deps.await?.len(),
+        };
+        std::fs::create_dir_all(crate_path.parent().unwrap())?;
+        let contents = ron::to_string(&res)?;
+        std::fs::write(crate_path, contents)?;
+        Ok(res)
+    }
 }
 
 async fn format_crates<T: AsRef<str>>(client: &AsyncClient, crates: &[T]) -> Result<()> {

### 2025-08-25 main.tex
fixup! cleanup
add more comments and notes for included crates

@@
     \item New fields of study (computational physics, chemistry, biology, sociology, etc.)
-    \item Cite some groundbreaking examples which were possible due to programming in science (maybe
-          weather)
+    \item Cite some groundbreaking examples which were possible due to programming in science (maybe weather)
     \item
@@
     \item modern features missing: package manager with package registries, build system
-    \item technical details; mutable by default, templates vs generics, backwards compatiblity
-          becomes a problem ("there is a much simpler langauge within C++ that tries to get out"),
-          macros, no clear ownership rules on a language level, dangling pointers
-    \item still many good principles (such as RAII) and design principles/concepts (cite gang of
-          four)
+    \item technical details; mutable by default, templates vs generics, backwards compatiblity becomes a
+          problem ("there is a much simpler langauge within C++ that tries to get out"), macros, no clear
+          ownership rules on a language level, dangling pointers
+    \item still many good principles (such as RAII) and design principles/concepts (cite gang of four)
 \end{enumerate}
@@
     \item Java aimed to replace C due to pointers
-    \item Go also aims to replace it, go is garbage-collected; thus cannot be as well-performing as
-          C++ or C
+    \item Go also aims to replace it, go is garbage-collected; thus cannot be as well-performing as C++ or C
     \item Python, Javascript for high-level abstractions ("as glue"); no performance howerver
@@
 \begin{enumerate}
-    \item safety guarantees; ensures that code which we write actually performs what we set out to
-          do
+    \item safety guarantees; ensures that code which we write actually performs what we set out to do
     \item no dangling pointers
@@
 \begin{enumerate}
-    \item Amazon (AWS)\\
-          AWS developed Firecracker, a microVM monitoring service powering AWS Lambda and AWS
-          Fargate.
-          Firecracker is built primarily in Rust to offer fast cold boot times and secure
-          workload isolation.
-    \item Rust-supported runtimes like AWS Lambda also allow developers to write serverless
-          functions using Rust.
-    \item Google\\
-          Fuchsia, Google's modern operating system, uses Rust extensively for system components
+    \item Amazon (AWS)\\ AWS developed Firecracker, a microVM monitoring service powering AWS Lambda and AWS
+          Fargate. Firecracker is built primarily in Rust to offer fast cold boot times and secure workload
+          isolation.
+    \item Rust-supported runtimes like AWS Lambda also allow developers to write serverless functions using
+          Rust.
+    \item Google\\ Fuchsia, Google's modern operating system, uses Rust extensively for system components
           such as drivers and services to enhance security and reliability.
-    \item Google (Android)\\
-          Android integration: Rust has been adopted within the Android Open Source Project (AOSP)
-          to improve memory safety in critical system components. DEV Community Litslink
-    \item Meta (Facebook)\\
-          Mononoke, Facebook’s large-scale source control backend for Mercurial, was written in
-          Rust to improve performance and reliability in its monorepo.
-    \item Meta (Facebook)\\
-          Facebook employs Rust in developing CLI tools, distributed systems, and cron jobs.
-          Engineers write millions of lines of Rust internally.
-    \item Apple\\
-          Apple’s Cloud Traffic team and backend infrastructure have been shifting from C to Rust,
-          with new functionality being primarily built in Rust. This change was driven by
-          performance, security, and developer efficiency goals. Reddit
-    \item Microsoft\\
-          Although not part of FAANG, it’s noteworthy—Microsoft is rewriting parts of the Windows
-          kernel and system components in Rust (e.g., DWriteCore, GDI Regions), prioritizing memory
-          safety.
+    \item Google (Android)\\ Android integration: Rust has been adopted within the Android Open Source
+          Project (AOSP) to improve memory safety in critical system components. DEV Community Litslink
+    \item Meta (Facebook)\\ Mononoke, Facebook’s large-scale source control backend for Mercurial, was
+          written in Rust to improve performance and reliability in its monorepo.
+    \item Meta (Facebook)\\ Facebook employs Rust in developing CLI tools, distributed systems, and cron
+          jobs. Engineers write millions of lines of Rust internally.
+    \item Apple\\ Apple’s Cloud Traffic team and backend infrastructure have been shifting from C to Rust,
+          with new functionality being primarily built in Rust. This change was driven by performance,
+          security, and developer efficiency goals. Reddit
+    \item Microsoft\\ Although not part of FAANG, it’s noteworthy—Microsoft is rewriting parts of the Windows
+          kernel and system components in Rust (e.g., DWriteCore, GDI Regions), prioritizing memory safety.
     \item Azure services-like Azure IoT Edge—also incorporate Rust.
-    \item Linux Kernel\\
-          Only other supported language except from C
+    \item Linux Kernel\\ Only other supported language except from C
 \end{enumerate}
@@
 implementations of clustering approaches. Intended for large datasets in scientific applications.
+\url{https://dotblocks.com/}
 
@@
 
-Diffsol\\
-A crate for solving differential equations.
+stochastic-rs\\
+stochastic-rs is a high-performance Rust library for simulating and analyzing
+stochastic processes. Designed for applications in quantitative finance, AI training, and
+statistical modeling, it provides efficient tools to generate synthetic data and analyze complex
+stochastic systems.
 
@@
 
-ploc\\
-Efficient point location with ploc.
-
 NPB-Rust\\
-NAS parallel benchmarks in Rust.
+NAS parallel benchmarks in Rust. \cite{martins2025npbrustnasparallelbenchmarks}
 
@@
 Cellular agent-based modeling from a clean slate.
+
+ORMATEX\\
+Oak Ridge MATrix EXponential tools - methods to compute the matrix exponential
+
+orbweaver-rs\\
+A fast R library for working with Nodes in a graph.
+
+triple-r\\
+Recycle, Reuse, Reduce. triple-r is a high-performance Rust library that provides
+wrappers around standard library collections to enable the reuse of their memory allocations. By
+recycling the underlying memory of collections like HashMap and Vec, triple-r helps reduce
+allocation overhead in performance-critical applications.
+
+faucet (maybe remove; does this fit the scope?)\\
+Scale, deploy and route Plumber APIs and Shiny
+applications with ease and efficiency.
+
+qoqo\\
+A toolkit to represent quantum circuits by HQS Quantum Simulations.
 
@@
 
-Juice\\
-Juice your simulations: what science can learn from game development.
-
 eVaiutilities\\
-Research data management post eVai.
+A data management software for the analysis of the eVai output files. It allows the
+analysis of the genomic variants further such as analyzing the multiple genomic annotated variants,
+reference and alternate allele, enabling coordinate search, coordinate search with specified
+variants and annotation search across a large number of population.
 
@@
 
-Stochastic\\
-Stochastic calculus in Rust.
+qhull-rs\\
+Qhull computes the convex hull, Delaunay triangulation, Voronoi diagram, halfspace
+intersection about a point, furthest-site Delaunay triangulation, and furthest-site Voronoi
+diagram.
 
-Exponential\\
-Exponential time integration for stiff systems in Rust.
+HULK (maybe out of scope?)\\
+The robot control program and associated tools of the RoboCup SPL team
+HULKs. \url{https://hulks.de/hulk/} "Rust’s safety and performance features benefit real-time
+robotics, and the challenges of integrating it into an ecosystem traditionally dominated by C++ and
+Python."
 
-Rust\\
-Rust for chip design algorithms and EDA software.
+burn\\
+Burn is a next generation Deep Learning Framework that doesn't compromise on flexibility,
+efficiency and portability. \url{https://github.com/tracel-ai/burn}\\
+Commercial
 
-Using\\
-Using Rust to mitigate breaking changes in scientific software.
-
-The\\
-The pleasure and pain of developing cross-platform CPU vector code in Rust.
+Diffsol\\
+A crate for solving differential equations.
 
@@
 
-Rust\\
-Rust is RAD and this is why.
+\paragraph{Scientific Computing in Rust 2024 talks}
 
-Diffsol\\
-A crate for solving differential equations.
-
-Pixi\\
-The missing companion to Cargo.
-
-ploc\\
-Efficient point location with ploc.
-
-\paragraph{Scientific Computing in Rust 2024 talks}
+oxidicom\\
+A high-performance DICOM receiver for the ChRIS backend (CUBE).
+\url{https://chrisproject.org/}
 
@@
 rlst\\
-The Rust Linear Solver toolbox is an in-development project for dense and sparse linear algebra routines in Rust.
+The Rust Linear Solver toolbox is an in-development project for dense and sparse linear
+algebra routines in Rust.
 
@@
 
-sophus-rs\\
-A fast electromagnetics library with Rust and Python.
+cfsem\\
+Quasi-steady electromagnetics including filamentized approximations, Biot-Savart, and
+Grad-Shafranov.\\
+\url{https://cfs.energy/} Commercial
 
@@
 
-Rootfinders\\
+Argmin\\
 Rootfinders for Rust.
@@
 
+Symbolica\\
+Symbolica is a blazing fast computer algebra system for Python and Rust, born of a need
+to push the boundaries of computations in science and enterprise.
+
 extendr\\
@@
 
-Rusting\\
-Rusting RSRS.
+rsrs (no crate, just github repo)\\
+An implementation of the randomised strong recursive
+skeletonisation
 
-Physics\\
-Physics simulations in Bevy.
-
-An\\
-An embedded memory cache for persistent memory access.
+bevy\\
+Bevy is a refreshingly simple data-driven game engine built in Rust.
 
 Polars\\
-Polars in the factory: when databases trump collections for modeling.
+Polars is a DataFrame library for Rust. It is based on Apache Arrow’s memory model. Apache
+Arrow provides very cache efficient columnar data structures and is becoming the defacto standard
+for columnar data.
 
@@
 
-Writing\\
-Writing maths software in Rust after 20 years of C++: graph isomorphism and groups.
-
-Productive\\
-Productive and reliable constitutive modeling in Rust.
-
-Efficient\\
-Efficient 1-bit matrix approximations.
-
-Developing\\
+Roxido\\
 Developing Rust-based R packages using the Roxido framework.
@@
 
-Back\\
+Biomechanics Foundation\\
 Back to basics: rebuilding computational biomechanics crate by crate.
+Many crates (maybe just cite one crate):
+\begin{itemize}
+    \item chiron - Graphical and command-line interface tools for Biomechanics Foundation
+    \item gcv\_spline - Generalized Cross-Validated Splines for Interpolation and Derivation in Pure Rust
+    \item onager - Featherstone rigid-body physics library built in Rust
+    \item c3dio - A c3d parser, writer and editor written in Rust.
+    \item bevy\_c3d - A .c3d asset loader plugin for the Bevy engine.
+    \item Mannequin - Methods for kinematics and rigging, and dynamics in biomechanics applications and
+          animation
+    \item hominem - Biomechanical muscle models for simulating human motion
+\end{itemize}
+\url{https://biomechanicsfoundation.org/}
+\url{https://github.com/biomechanics-foundation}
 
-Cargo\\
-Cargo +GPU build: An early outlook.
+\#[offload] (feature draft)\\
+New rustc feature called \#[offload], which allows running almost arbitrary Rust code on
+co-processors like a GPU, TPU, or IPU
+\url{https://github.com/rust-lang/rust/pull/145768}
 
-Writing\\
-Writing a grid library for finite and boundary element methods.
+Bempp-rs\\
+Writing a grid library for finite and boundary element methods. \cite{Witherden2015}
 
@@
 alg\_tools\\
-utility routines and tools for implementing iterative algorithms and (abstract) numerical computing in Rust~\cite{Valkonen2023}.
-\url{https://tuomov.iki.fi/software/alg_tools/}
+utility routines and tools for implementing iterative algorithms and (abstract)
+numerical computing in Rust~\cite{Valkonen2023}. \url{https://tuomov.iki.fi/software/alg_tools/}
 
 clarabel\\
-Clarabel is an interior point numerical solver for convex optimization problems using a novel
-homogeneous embedding~\cite{Clarabel_2024}.
+Clarabel is an interior point numerical solver for convex optimization problems using a
+novel homogeneous embedding~\cite{Clarabel_2024}.
 
@@
 
-\paragraph{FAANG Companies}
-
 \end{document}

### 2025-08-25 main.tex
deduplicate list of crates

@@
 
-EGObox\\
-A library for evolutionary game theory and agent-based modeling. It allows researchers to
-explore strategies, equilibria, and dynamics in multi-agent systems. Designed with flexibility to
-model diverse game-theoretic scenarios.
-
 faer\\
@@
 
-kifmm-rs\\
-A Rust implementation of the kernel-independent fast multipole method (FMM). It
-accelerates particle simulations and large-scale interactions in physics. The focus is on
-high-performance computing and scalability.
-
 ndgrid\\
@@
 \url{https://dotblocks.com/}
-
-Polars\\
-A fast DataFrame library for Rust, with APIs also available in Python. It provides
-powerful data manipulation, joins, and group-by operations with performance close to Apache Arrow.
-Designed for data science, analytics, and scientific workflows.
 
@@
 \url{https://chrisproject.org/}
+ChRIS is an open-source platform for computational research and medicine.
 
@@
 algebra routines in Rust.
-
-Raphtory\\
-Temporal graphs from Scala to Rust.
 
@@
 
-extendr\\
-Frictionless bindings for R and Rust.
-
 rsrs (no crate, just github repo)\\
@@
 Arrow provides very cache efficient columnar data structures and is becoming the defacto standard
-for columnar data.
+for columnar data.\\
+A fast DataFrame library for Rust, with APIs also available in Python. It provides
+powerful data manipulation, joins, and group-by operations with performance close to Apache Arrow.
+Designed for data science, analytics, and scientific workflows.
 
@@
 Perpetaul: a hyperparameter-free gradient boosting machine.
-
-Roxido\\
-Developing Rust-based R packages using the Roxido framework.
-
-Rusph\\
-Rusph: a SPH astrophysical simulation code in the Rust programming language.
 
@@
 
-\url{https://chrisproject.org/}\\
-ChRIS is an open-source platform for computational research and medicine.
-
 struqture\\
@@
 R extensions written in Rust
-
-faer-rs\\
-A linear algebra foundation for the Rust programming language.
 
@@
 
-Rstats\\
-Multidimensional data analysis in Rust.
-
-extendr\\
-Frictionless bindings for R and Rust.
-
-Rusting\\
-Rusting RSRS.
-
-Physics\\
-Physics simulations in Bevy.
-
-An\\
-An embedded memory cache for persistent memory access.
-
-Polars\\
-Polars in the factory: when databases trump collections for modeling.
-
-Perpetaul\\
-Perpetaul: a hyperparameter-free gradient boosting machine.
-
-Writing\\
-Writing maths software in Rust after 20 years of C++: graph isomorphism and groups.
-
-Productive\\
-Productive and reliable constitutive modeling in Rust.
-
-Efficient\\
-Efficient 1-bit matrix approximations.
-
-Developing\\
-Developing Rust-based R packages using the Roxido framework.
-
 Rusph\\
@@
 
-Back\\
-Back to basics: rebuilding computational biomechanics crate by crate.
-
-Cargo\ \ Cargo +GPU build: An early outlook.
-
-Writing\\
-Writing a grid library for finite and boundary element methods.
+Roxido\\
+Developing Rust-based R packages using the Roxido framework.
 

### 2025-08-26 crate-info-puller/src/lib.rs
exttract functionality to lib.rs and binary

@@ -1,0 +1,54 @@
+use anyhow::Result;
+use crates_io_api_wasm_patch::{AsyncClient, CrateDownloads, CrateResponse};
+use serde::{Deserialize, Serialize};
+
+#[derive(Clone, Deserialize, Serialize, Debug)]
+pub struct CrateData {
+    pub crate_name: String,
+    pub crate_response: CrateResponse,
+    pub downloads: CrateDownloads,
+    pub deps: usize,
+}
+
+pub async fn get_data(client: &AsyncClient, crate_name: &str) -> Result<CrateData> {
+    // First check if data was stored to disk
+    let mut crate_path = std::path::PathBuf::new();
+    crate_path.push("./out/");
+    crate_path.push(crate_name);
+
+    if crate_path.exists() {
+        let content = std::fs::read_to_string(crate_path)?;
+        let res: CrateData = ron::from_str(&content)?;
+        Ok(res)
+    } else {
+        let dow = client.crate_downloads(crate_name);
+        let crate_response = client.get_crate(crate_name).await?;
+        let num = &crate_response.versions.first().unwrap().num.clone();
+        let deps = client.crate_dependencies(crate_name, num);
+
+        let res = CrateData {
+            crate_name: crate_name.to_string(),
+            crate_response,
+            downloads: dow.await?,
+            deps: deps.await?.len(),
+        };
+        std::fs::create_dir_all(crate_path.parent().unwrap())?;
+        let contents = ron::to_string(&res)?;
+        std::fs::write(crate_path, contents)?;
+        Ok(res)
+    }
+}
+
+/// Query crates.io and generate a summary table
+///
+/// The crate names are sorted by name before any output is produced.
+#[derive(clap::Parser, Debug)]
+#[command(version, about, max_term_width = 60)]
+pub struct Args {
+    /// Comma-separated list of crate names
+    #[arg(short, long, value_delimiter = ',')]
+    pub crates: Vec<String>,
+    /// Read crate names from lines in a text file
+    #[arg(short, long)]
+    pub file: Option<String>,
+}

### 2025-08-26 crate-info-puller/src/tabular.rs

@@ -1,45 +1,9 @@
 use anyhow::{Context, Result};
-use crates_io_api_wasm_patch::{AsyncClient, CrateDownloads, CrateResponse};
-use serde::{Deserialize, Serialize};
+use crates_io_api_wasm_patch::AsyncClient;
 
-#[derive(Clone, Deserialize, Serialize, Debug)]
-struct CrateData {
-    crate_name: String,
-    crate_response: CrateResponse,
-    downloads: CrateDownloads,
-    deps: usize,
-}
+use crate_info_puller::*;
 
-async fn get_data(client: &AsyncClient, crate_name: &str) -> Result<CrateData> {
-    // First check if data was stored to disk
-    let mut crate_path = std::path::PathBuf::new();
-    crate_path.push("./out/");
-    crate_path.push(crate_name);
-
-    if crate_path.exists() {
-        let content = std::fs::read_to_string(crate_path)?;
-        let res: CrateData = ron::from_str(&content)?;
-        Ok(res)
-    } else {
-        let dow = client.crate_downloads(crate_name);
-        let crate_response = client.get_crate(crate_name).await?;
-        let num = &crate_response.versions.first().unwrap().num.clone();
-        let deps = client.crate_dependencies(crate_name, num);
-
-        let res = CrateData {
-            crate_name: crate_name.to_string(),
-            crate_response,
-            downloads: dow.await?,
-            deps: deps.await?.len(),
-        };
-        std::fs::create_dir_all(crate_path.parent().unwrap())?;
-        let contents = ron::to_string(&res)?;
-        std::fs::write(crate_path, contents)?;
-        Ok(res)
-    }
-}
-
-async fn format_crates<T: AsRef<str>>(client: &AsyncClient, crates: &[T]) -> Result<()> {
+pub async fn format_crates<T: AsRef<str>>(client: &AsyncClient, crates: &[T]) -> Result<()> {
     println!("\\begin{{tabular}}{{l r r}}");
     println!("    Crate Name &Weekly Downloads &Last Update &Latest Version &Dependencies\\\\");
     for (n, name) in crates.iter().enumerate() {
@@ -74,7 +38,7 @@ async fn format_crates<T: AsRef<str>>(client: &AsyncClient, crates: &[T]) -> Res
         print!(
             "    {:16} &{weekly_downloads:<8.0} &{:10} &{latest_version:10} &{n_deps}",
             cd.crate_name,
-            last_update.format("%d/%m/%Y")
[…]

### 2025-08-26 crate-info-puller/src/growth.rs
new binary to extract growth data from crates.io

@@ -1,0 +1,76 @@
+use anyhow::Result;
+use chrono::Datelike;
+use crates_io_api_wasm_patch::AsyncClient;
+
+use crate_info_puller::*;
+
+pub async fn format_crates<T: AsRef<str>>(client: &AsyncClient, crates: &[T]) -> Result<()> {
+    let mut all_downloads = std::collections::BTreeMap::<chrono::DateTime<chrono::Utc>, u64>::new();
+    for name in crates.iter() {
+        let cd = async_compat::Compat::new(get_data(client, name.as_ref())).await?;
+
+        for version in cd.crate_response.versions.into_iter() {
+            let date = version.updated_at;
+            // let date = updated_at.format("%d/%m/%Y").to_string();
+            let entry = all_downloads.entry(date).or_insert(0);
+            *entry += 1;
+        }
+    }
+
+    let tmin = all_downloads.keys().min().unwrap();
+    let tmax = all_downloads.keys().max().unwrap();
+
+    let interval = chrono::Duration::weeks(4);
+
+    let mut skip_size = 0;
+    let mut t = *tmin;
+    let mut output = std::collections::BTreeMap::new();
+    while &t <= tmax {
+        t += interval;
+        let n = all_downloads
+            .iter()
+            .skip(skip_size)
+            .take_while(|(d, _)| d < &&t)
+            .count();
+
+        skip_size += n;
+
+        use anyhow::Context;
+        if let Some(x) = output.insert(t, n) {
+            None.context(format!("Should not contain any value at this point: {x}"))?;
+        }
+    }
+
+    for (month, counts) in output.iter() {
+        println!("{}/{},{counts}", month.month(), month.year());
+    }
+
+    Ok(())
+}
+
+#[macro_rules_attribute::apply(smol_macros::main)]
+async fn main() -> Result<()> {
+    use clap::Parser;
+    use reqwest::header::*;
+
+    let args = Args::parse();
+    let mut crates = args.crates;
+    if let Some(filename) = args.file {
+        let content = std::fs::read_to_string(filename)?;
+        crates.extend(content.lines().map(String::from));
+    }
+    crates.sort();
+
+    let mut headers = HeaderMap::new();
+    headers.insert(USER_AGENT, HeaderValue::from_str("gobbler")?);
+
+    let client1 = reqwest::Client::builder()
+        .default_headers(headers)
+        .build()?;
+    let client =
[…]

### 2025-08-26 crate-info-puller/src/growth.rs
directly write to file

@@ -41,8 +41,10 @@ pub async fn format_crates<T: AsRef<str>>(client: &AsyncClient, crates: &[T]) ->
         }
     }
 
+    let mut file = std::fs::File::create("./crate-lists/releases.csv")?;
     for (month, counts) in output.iter() {
-        println!("{}/{},{counts}", month.month(), month.year());
+        use std::io::Write;
+        writeln!(&mut file, "{}/{},{counts}", month.month(), month.year())?;
     }
 
     Ok(())

### 2025-08-26 figures/crates-io-release-numbers.py
new plot with output files

@@ -1,0 +1,61 @@
+import csv
+import importlib
+import scipy as sp
+
+pypi_mod = importlib.import_module("pypi-org-used-languages", package=None)
+
+if __name__ == "__main__":
+    res = []
+    with open("crate-info-puller/crate-lists/releases.csv", newline="") as csvfile:
+        reader = csv.reader(csvfile, delimiter=",")
+        for n, row in enumerate(reader):
+            if n != 0:
+                res.append(row)
+
+    x = [r[0] for r in res]
+    y = [int(r[1]) for r in res]
+
+    fig, ax = pypi_mod.prepare_fig_ax()
+
+    ax.plot(x[1:], y[1:], color="#CE412B", label="Monthly Releases")
+
+    ticks = ax.get_xticks()
+    labels = ax.get_xticklabels()
+
+    newticks = []
+    newlabels = []
+    for tick, label in zip(ticks, labels):
+        split = label.get_text().split("/")
+        if split[0] == "1":
+            newticks.append(tick)
+            newlabels.append(split[1])
+
+    # Do linear fit
+    x_fit = list(range(len(x) - 1))
+    popt, pcov = sp.optimize.curve_fit(
+        lambda x, a, b, c: a * x**2 + b * x + c, x_fit, y[1:], p0=(1, 0, 0)
+    )
+    a, b, c = popt
+    da = pcov[0][0] ** 0.5
+    db = pcov[1][1] ** 0.5
+    dc = pcov[2][2] ** 0.5
+
+    y_fit = [a * xi**2 + b * xi + c for xi in x_fit]
+    # y_fit_lower = [(a - da) * xi**2 + (b - db) * xi + c - dc for xi in x_fit]
+    # y_fit_upper = [(a + da) * xi**2 + (b + db) * xi + c + dc for xi in x_fit]
+
+    ax.plot(x[1:], y_fit, color="#00599d", linestyle="--", label="Quadratic Fit")
+    # ax.fill_between(x[1:], y_fit_lower, y_fit_upper, color="#00599d", alpha=0.2)
+
+    ax.set_xticks(newticks, newlabels, rotation=45)
+
+    ax.legend(
+        ncol=2,
+        frameon=False,
+        loc="upper center",
+        bbox_to_anchor=(0.5, 1.15),
+    )
+
+    fig.savefig("figures/crates-io-release-numbers.png")
+    fig.savefig("figures/crates-io-release-numbers.pdf")
+    fig.savefig("figures/crates-io-release-numbers.svg")

### 2025-08-30 figures/crates-io-release-numbers.py
only plot every 2nd year and without angle

@@ -26,7 +26,7 @@ if __name__ == "__main__":
     newlabels = []
     for tick, label in zip(ticks, labels):
         split = label.get_text().split("/")
-        if split[0] == "1":
+        if split[0] == "1" and int(split[1]) % 2 == 1:
             newticks.append(tick)
             newlabels.append(split[1])
 
@@ -47,7 +47,7 @@ if __name__ == "__main__":
     ax.plot(x[1:], y_fit, color="#00599d", linestyle="--", label="Quadratic Fit")
     # ax.fill_between(x[1:], y_fit_lower, y_fit_upper, color="#00599d", alpha=0.2)
 
-    ax.set_xticks(newticks, newlabels, rotation=45)
+    ax.set_xticks(newticks, newlabels)
 
     ax.legend(
         ncol=2,

### 2025-08-30 paper/main.tex
move paper to own subfolder

--- 2025-08-25 main.tex
@@
 \tableofcontents
+
+\twocolumn
 \pagebreak
@@
 
-\begin{figure}
+\begin{figure*}
     \centering
@@
     \includegraphics[width=0.5\textwidth]{figures/pypi-org-used-languages.pdf}
+    \caption{
+        Usage of programming languages according to (A) stackoverflow.com and (B) pypi.org.
+    }
+\end{figure*}
+\begin{figure*}
+    \centering
     \includegraphics[width=0.5\textwidth]{figures/stackoverflow-loved-language.pdf}%
     \includegraphics[width=0.5\textwidth]{figures/stackoverflow-desired-language.pdf}
-    \caption{TODO}
-\end{figure}
+    \caption{
+        Popularity of the Rust programming language.
+    }
+\end{figure*}
+\begin{figure*}
+    \centering
+    \includegraphics[width=0.5\textwidth]{figures/crates-io-release-numbers.pdf}
+    \caption{
+        Number of releases of all included crates.
+        Measure of overall growth of the field.
+    }
+\end{figure*}
 
@@
 \subsection{Foundational Components}
+Definition:\\
+Lowest level of library considered in this review. Other crates depend on these
+foundational crates.
+
 \subsubsection{Linear Algebra \& Arrays}
@@
 \subsection{Computational Methods}
+Definition:\\
+They are typically built on top of Foundational Components. Can also generalize over
+many of these (such as different linear algebra crates).
+
 \subsubsection{Optimization}
@@
 \subsection{High-Performance \& Scalable Computing}
+Definition:\\
+Frameworks/libraries that allow greater scalability or usage of additional resources
+such as GPUs/TPUs etc.
+
 \subsubsection{Parallel Computing}
@@
 \subsection{Creature Comforts}
+Definition:\\
+Everything that is still missing
+
 \subsubsection{Debugging}
@@
 \subsection{Visualization}
+Definition:\\
+Libraries which can be used to visualize results. (self-explanatory)
+
 \subsubsection{Plotting (2D)}
@@
 
-Meshing (honeycomb)
+Meshing (honeycomb) (maybe this should be in section "Computational Methods" / "Differential
+Equations")
 
@@
 
+\begin{itemize}
+    \item Job Opportunities remain a problem in Rust
+\end{itemize}
+
 %###################################################################################################
@@
 
+\onecolumn
 \printbibliography
@@
 
-qoqo\\
+qoqo (Roqoqo)\\
 A toolkit to represent quantum circuits by HQS Quantum Simulations.
@@
 qhull-rs\\
-Qhull computes the convex hull, Delaunay triangulation, Voronoi diagram, halfspace
-intersection about a point, furthest-site Delaunay triangulation, and furthest-site Voronoi
-diagram.
+Safe Rust Qhull bindings. Qhull computes the convex hull, Delaunay triangulation,
+Voronoi diagram, halfspace intersection about a point, furthest-site Delaunay triangulation, and
+furthest-site Voronoi diagram.
 
@@
 A high-performance DICOM receiver for the ChRIS backend (CUBE).
-\url{https://chrisproject.org/}
-ChRIS is an open-source platform for computational research and medicine.
+\url{https://chrisproject.org/} ChRIS is an open-source platform for computational research and
+medicine.
 
@@
 for columnar data.\\
-A fast DataFrame library for Rust, with APIs also available in Python. It provides
-powerful data manipulation, joins, and group-by operations with performance close to Apache Arrow.
-Designed for data science, analytics, and scientific workflows.
+A fast DataFrame library for Rust, with APIs also available in Python. It
+provides powerful data manipulation, joins, and group-by operations with performance close to
+Apache Arrow. Designed for data science, analytics, and scientific workflows.
 
@@
 
-Roqoqo\\
-A quantum computing toolkit in Rust.
-
 Lace\\

### 2025-08-30 figures/crates-io-release-numbers.py
store under new path

@@ -56,6 +56,6 @@ if __name__ == "__main__":
         bbox_to_anchor=(0.5, 1.15),
     )
 
-    fig.savefig("figures/crates-io-release-numbers.png")
-    fig.savefig("figures/crates-io-release-numbers.pdf")
-    fig.savefig("figures/crates-io-release-numbers.svg")
+    fig.savefig("paper/figures/crates-io-release-numbers.png")
+    fig.savefig("paper/figures/crates-io-release-numbers.pdf")
+    fig.savefig("paper/figures/crates-io-release-numbers.svg")

### 2025-08-30 figures/pypi-org-used-languages.py

@@ -48,6 +48,6 @@ ax.barh(
 ax.set_xlabel("(B) Number of Packages")
 ax.set_title("Packages at pypi.org")
 
-fig.savefig("figures/pypi-org-used-languages.png")
-fig.savefig("figures/pypi-org-used-languages.pdf")
-fig.savefig("figures/pypi-org-used-languages.svg")
+fig.savefig("paper/figures/pypi-org-used-languages.png")
+fig.savefig("paper/figures/pypi-org-used-languages.pdf")
+fig.savefig("paper/figures/pypi-org-used-languages.svg")

### 2025-08-30 figures/stackoverflow-desired-language.py

@@ -58,10 +58,10 @@ for n, ax in enumerate([ax1, ax2]):
         bbox_to_anchor=(1.17, 0.5),
     )
 
-fig1.savefig("figures/stackoverflow-loved-language.png")
-fig1.savefig("figures/stackoverflow-loved-language.pdf")
-fig1.savefig("figures/stackoverflow-loved-language.svg")
+fig1.savefig("paper/figures/stackoverflow-loved-language.png")
+fig1.savefig("paper/figures/stackoverflow-loved-language.pdf")
+fig1.savefig("paper/figures/stackoverflow-loved-language.svg")
 
-fig2.savefig("figures/stackoverflow-desired-language.png")
-fig2.savefig("figures/stackoverflow-desired-language.pdf")
-fig2.savefig("figures/stackoverflow-desired-language.svg")
+fig2.savefig("paper/figures/stackoverflow-desired-language.png")
+fig2.savefig("paper/figures/stackoverflow-desired-language.pdf")
+fig2.savefig("paper/figures/stackoverflow-desired-language.svg")

### 2025-08-30 figures/stackoverflow-popular-languages.py

@@ -57,6 +57,6 @@ ax.legend(
 # ax.set_title("Language Popularity (stackoverflow.co)")
 ax.set_xlabel("(A) Language Usage [%]")
 
-fig.savefig("figures/stackoverflow-popular-languages.png")
-fig.savefig("figures/stackoverflow-popular-languages.pdf")
-fig.savefig("figures/stackoverflow-popular-languages.svg")
+fig.savefig("paper/figures/stackoverflow-popular-languages.png")
+fig.savefig("paper/figures/stackoverflow-popular-languages.pdf")
+fig.savefig("paper/figures/stackoverflow-popular-languages.svg")

### 2025-09-01 crate-info-puller/src/tabular.rs
correctly filter crate versions

@@ -13,18 +13,39 @@ pub async fn format_crates<T: AsRef<str>>(client: &AsyncClient, crates: &[T]) ->
         d.sort_by_key(|x| x.date);
 
         let last_update = cd.crate_response.crate_data.updated_at;
-        let latest_version = cd
+
+        let (_, &latest_version_id, latest_version) = cd
             .crate_response
             .crate_data
             .versions
-            .and_then(|v| v.into_iter().max())
-            .context("no version provided by crates.io")?;
+            .iter()
+            .flatten()
+            .zip(cd.crate_response.versions.iter())
+            .filter_map(|(id, version)| {
+                if !version.yanked
+                    && !version.num.contains("dev")
+                    && !version.num.contains("alpha")
+                    && !version.num.contains("rc")
+                {
+                    semver::Version::parse(&version.num)
+                        .ok()
+                        .map(|x| (x, id, version))
+                } else {
+                    None
+                }
+            })
+            .max_by_key(|(smvr, _, _)| smvr.clone())
+            .unwrap_or((
+                semver::Version::parse(&cd.crate_response.versions[0].num)?,
+                &cd.crate_response.crate_data.versions.as_ref().unwrap()[0],
+                &cd.crate_response.versions[0],
+            ));
 
         let weekly_downloads = {
             let mut n = 0;
             let w = d
                 .into_iter()
-                .filter(|x| x.version == latest_version)
+                .filter(|x| x.version == latest_version_id)
                 .map(|di| {
                     n += 1;
                     di.downloads

### 2025-09-01 crate-info-puller/src/tabular.rs
change formatting of tabular

@@ -4,8 +4,12 @@ use crates_io_api_wasm_patch::AsyncClient;
 use crate_info_puller::*;
 
 pub async fn format_crates<T: AsRef<str>>(client: &AsyncClient, crates: &[T]) -> Result<()> {
-    println!("\\begin{{tabular}}{{l r r}}");
-    println!("    Crate Name &Weekly Downloads &Last Update &Latest Version &Dependencies\\\\");
+    println!("\\begin{{tabular}}{{l r r r r r r}}");
+    println!("    \\toprule");
+    println!(
+        "    Crate Name &Last Update &Version &$N_d$ &$D_\\text{{week}}$ &$D_\\text{{total}}$\\\\"
+    );
+    println!("    \\midrule");
     for (n, name) in crates.iter().enumerate() {
         let cd = async_compat::Compat::new(get_data(client, name.as_ref())).await?;
 
@@ -53,13 +57,18 @@ pub async fn format_crates<T: AsRef<str>>(client: &AsyncClient, crates: &[T]) ->
                 .sum::<u64>();
             w as f64 / n as f64 * 7.
         };
-        let latest_version = cd.crate_response.versions[0].num.clone();
+        let total_downloads = cd.crate_response.crate_data.downloads;
         let n_deps = cd.deps;
 
+        let tot = format!("{:>7.1}k", total_downloads as f64 / 1_000.);
         print!(
-            "    {:16} &{weekly_downloads:<8.0} &{:10} &{latest_version:10} &{n_deps}",
+            "    {:16} &{:10} &{:10} &{:<4.0} &{:<8.0} &{:<8}",
             cd.crate_name,
-            last_update.format("%d-%m-%Y")
+            last_update.format("%Y-%m-%d"),
+            latest_version.num,
+            n_deps,
+            weekly_downloads,
+            tot,
         );
         if n + 1 < crates.len() {
             println!("\\\\");
@@ -67,6 +76,7 @@ pub async fn format_crates<T: AsRef<str>>(client: &AsyncClient, crates: &[T]) ->
             println!();
         }
     }
+    println!("    \\bottomrule");
     println!("\\end{{tabular}}");
     Ok(())
 }

### 2025-09-01 crate-info-puller/src/tabular.rs
cleanup

@@ -1,4 +1,4 @@
-use anyhow::{Context, Result};
+use anyhow::Result;
 use crates_io_api_wasm_patch::AsyncClient;
 
 use crate_info_puller::*;
@@ -12,9 +12,6 @@ pub async fn format_crates<T: AsRef<str>>(client: &AsyncClient, crates: &[T]) ->
     println!("    \\midrule");
     for (n, name) in crates.iter().enumerate() {
         let cd = async_compat::Compat::new(get_data(client, name.as_ref())).await?;
-
-        let mut d = cd.downloads.version_downloads;
-        d.sort_by_key(|x| x.date);
 
         let last_update = cd.crate_response.crate_data.updated_at;
 
@@ -47,7 +44,9 @@ pub async fn format_crates<T: AsRef<str>>(client: &AsyncClient, crates: &[T]) ->
 
         let weekly_downloads = {
             let mut n = 0;
-            let w = d
+            let w = cd
+                .downloads
+                .version_downloads
                 .into_iter()
                 .filter(|x| x.version == latest_version_id)
                 .map(|di| {

### 2025-09-01 paper/main.tex
formatting
cleanup
example table for crates

@@
 \usepackage{doi}
-\usepackage{acronym}
 \usepackage{listings}
@@
 \usetikzlibrary{trees}
-
-\newacro{abm}[ABM]{Agent-Based Model}
-\newacro{cabm}[CABM]{Cellular Agent-Based Model}
-\newacro{ca}[CA]{Cellular Automaton}
-\newacroplural{ca}[CA]{Cellular Automata}
 
@@
 
+\begin{table*}
+    \centering
+    \begin{tabular}{l r r r r r}
+        \toprule
+        Crate Name &Last Update &Version &$N_d$ &$D_\text{week}$ &$D_\text{total}$\\
+        \midrule
+        Lamellar         &2025-03-31 &0.7.1      &42   &20       &     16.9k\\
+        argmin           &2024-02-27 &0.10.0     &18   &12003    &   1227.7k\\
+        arrow            &2025-08-25 &56.1.0     &23   &7532     &  27425.0k\\
+        bevy             &2025-05-30 &0.16.1     &35   &10175    &   2847.7k\\
+        bless            &2025-07-20 &0.1.1      &13   &80       &      0.6k\\
+        burn             &2025-07-18 &0.18.0     &12   &1793     &    319.7k\\
+        c3dio            &2024-04-08 &0.8.0      &2    &24       &     11.8k\\
+        cargo-upgrades   &2025-08-25 &2.2.4      &10   &553      &     64.7k\\
+        cellular\_raza   &2025-08-05 &0.3.0      &10   &157      &     27.4k\\
+        cfsem            &2025-08-14 &2.5.0      &8    &106      &      4.9k\\
+        city2ba          &2020-02-28 &1.1.0      &20   &19       &      3.0k\\
+        cmake            &2025-02-10 &0.1.54     &1    &899615   &  91211.9k\\
+        cudarc           &2025-08-07 &0.17.2     &5    &1184     &    647.0k\\
+        cxxbridge        &2019-12-28 &0.0.0      &0    &14       &      1.3k\\
+        deimos           &2025-06-23 &0.10.0     &13   &40       &      7.1k\\
+        diffsol          &2025-07-16 &0.6.6      &16   &131      &     33.9k\\
+        dioxus           &2025-08-11 &0.6.3      &28   &         &    637.5k\\
+        egobox           &2025-08-22 &0.32.0     &18   &250      &     35.6k\\
+        enzyme           &2021-12-26 &0.4.0      &5    &17       &      5.4k\\
+        extendr-api      &2025-07-12 &0.8.1      &11   &968      &    178.4k\\
+        factrs           &2025-01-13 &0.2.0      &20   &24       &      1.8k\\
+        faer             &2025-04-29 &0.22.6     &46   &1322     &    528.4k\\
+        faucet           &2025-05-31 &0.2.0      &7    &33       &      3.0k\\
+        fem\_2D          &2023-04-20 &0.2.2      &6    &20       &      5.4k\\
+        gcv\_spline      &2024-01-28 &0.2.0      &1    &16       &      2.5k\\
+        \bottomrule
+    \end{tabular}
+    \caption{TODO}
+\end{table*}
+
 General Remarks about languages
@@
     \item New fields of study (computational physics, chemistry, biology, sociology, etc.)
-    \item Cite some groundbreaking examples which were possible due to programming in science (maybe weather)
+    \item Cite some groundbreaking examples which were possible due to programming in
+        science (maybe weather)
     \item
@@
     \item modern features missing: package manager with package registries, build system
-    \item technical details; mutable by default, templates vs generics, backwards compatiblity becomes a
-          problem ("there is a much simpler langauge within C++ that tries to get out"), macros, no clear
+    \item technical details; mutable by default, templates vs generics, backwards
+        compatiblity becomes a
+        problem ("there is a much simpler langauge within C++ that tries to get out"),
+        macros, no clear
         ownership rules on a language level, dangling pointers
-    \item still many good principles (such as RAII) and design principles/concepts (cite gang of four)
+    \item still many good principles (such as RAII) and design principles/concepts (cite
+        gang of four)
 \end{enumerate}
@@
     \item Java aimed to replace C due to pointers
-    \item Go also aims to replace it, go is garbage-collected; thus cannot be as well-performing as C++ or C
+    \item Go also aims to replace it, go is garbage-collected; thus cannot be as
+        well-performing as C++ or C
     \item Python, Javascript for high-level abstractions ("as glue"); no performance howerver
@@
 \begin{enumerate}
-    \item safety guarantees; ensures that code which we write actually performs what we set out to do
+    \item safety guarantees; ensures that code which we write actually performs what we
+        set out to do
     \item no dangling pointers
@@
 \begin{enumerate}
-    \item Amazon (AWS)\\ AWS developed Firecracker, a microVM monitoring service powering AWS Lambda and AWS
-          Fargate. Firecracker is built primarily in Rust to offer fast cold boot times and secure workload
+    \item Amazon (AWS)\\ AWS developed Firecracker, a microVM monitoring service powering
+        AWS Lambda and AWS
+        Fargate. Firecracker is built primarily in Rust to offer fast cold boot times and
+        secure workload
         isolation.
-    \item Rust-supported runtimes like AWS Lambda also allow developers to write serverless functions using
+    \item Rust-supported runtimes like AWS Lambda also allow developers to write
+        serverless functions using
         Rust.
-    \item Google\\ Fuchsia, Google's modern operating system, uses Rust extensively for system components
+    \item Google\\ Fuchsia, Google's modern operating system, uses Rust extensively for
+        system components
         such as drivers and services to enhance security and reliability.
-    \item Google (Android)\\ Android integration: Rust has been adopted within the Android Open Source
-          Project (AOSP) to improve memory safety in critical system components. DEV Community Litslink
-    \item Meta (Facebook)\\ Mononoke, Facebook’s large-scale source control backend for Mercurial, was
+    \item Google (Android)\\ Android integration: Rust has been adopted within the
+        Android Open Source
+        Project (AOSP) to improve memory safety in critical system components. DEV
+        Community Litslink
+    \item Meta (Facebook)\\ Mononoke, Facebook’s large-scale source control backend for
+        Mercurial, was
         written in Rust to improve performance and reliability in its monorepo.
-    \item Meta (Facebook)\\ Facebook employs Rust in developing CLI tools, distributed systems, and cron
+    \item Meta (Facebook)\\ Facebook employs Rust in developing CLI tools, distributed
+        systems, and cron
         jobs. Engineers write millions of lines of Rust internally.
-    \item Apple\\ Apple’s Cloud Traffic team and backend infrastructure have been shifting from C to Rust,
+    \item Apple\\ Apple’s Cloud Traffic team and backend infrastructure have been
+        shifting from C to Rust,
         with new functionality being primarily built in Rust. This change was driven by performance,
         security, and developer efficiency goals. Reddit
-    \item Microsoft\\ Although not part of FAANG, it’s noteworthy—Microsoft is rewriting parts of the Windows
-          kernel and system components in Rust (e.g., DWriteCore, GDI Regions), prioritizing memory safety.
+    \item Microsoft\\ Although not part of FAANG, it’s noteworthy—Microsoft is rewriting
+        parts of the Windows
+        kernel and system components in Rust (e.g., DWriteCore, GDI Regions),
+        prioritizing memory safety.
     \item Azure services-like Azure IoT Edge—also incorporate Rust.
@@
     \item chiron - Graphical and command-line interface tools for Biomechanics Foundation
-    \item gcv\_spline - Generalized Cross-Validated Splines for Interpolation and Derivation in Pure Rust
+    \item gcv\_spline - Generalized Cross-Validated Splines for Interpolation and
+        Derivation in Pure Rust
     \item onager - Featherstone rigid-body physics library built in Rust
@@
     \item bevy\_c3d - A .c3d asset loader plugin for the Bevy engine.
-    \item Mannequin - Methods for kinematics and rigging, and dynamics in biomechanics applications and
+    \item Mannequin - Methods for kinematics and rigging, and dynamics in biomechanics
+        applications and
         animation

### 2025-09-01 crate-info-puller/src/tabular.rs
use correct number of cols

@@ -4,7 +4,7 @@ use crates_io_api_wasm_patch::AsyncClient;
 use crate_info_puller::*;
 
 pub async fn format_crates<T: AsRef<str>>(client: &AsyncClient, crates: &[T]) -> Result<()> {
-    println!("\\begin{{tabular}}{{l r r r r r r}}");
+    println!("\\begin{{tabular}}{{l r r r r r}}");
     println!("    \\toprule");
     println!(
         "    Crate Name &Last Update &Version &$N_d$ &$D_\\text{{week}}$ &$D_\\text{{total}}$\\\\"

### 2025-09-01 crate-info-puller/src/tabular.rs
do not omit linebreak at end of table
- this was previously included for tables which did not have a
  bottomrule

@@ -10,7 +10,7 @@ pub async fn format_crates<T: AsRef<str>>(client: &AsyncClient, crates: &[T]) ->
         "    Crate Name &Last Update &Version &$N_d$ &$D_\\text{{week}}$ &$D_\\text{{total}}$\\\\"
     );
     println!("    \\midrule");
-    for (n, name) in crates.iter().enumerate() {
+    for name in crates.iter() {
         let cd = async_compat::Compat::new(get_data(client, name.as_ref())).await?;
 
         let last_update = cd.crate_response.crate_data.updated_at;
@@ -60,20 +60,19 @@ pub async fn format_crates<T: AsRef<str>>(client: &AsyncClient, crates: &[T]) ->
         let n_deps = cd.deps;
 
         let tot = format!("{:>7.1}k", total_downloads as f64 / 1_000.);
-        print!(
-            "    {:16} &{:10} &{:10} &{:<4.0} &{:<8.0} &{:<8}",
-            cd.crate_name,
+        println!(
+            "    {:16} &{:10} &{:10} &{:<4.0} &{:<8} &{:>10}\\\\",
+            cd.crate_name.replace("_", "\\_"),
             last_update.format("%Y-%m-%d"),
             latest_version.num,
             n_deps,
-            weekly_downloads,
+            if weekly_downloads.is_nan() {
+                String::new()
+            } else {
+                format!("{weekly_downloads:<8.0}")
+            },
             tot,
         );
-        if n + 1 < crates.len() {
-            println!("\\\\");
-        } else {
-            println!();
-        }
     }
     println!("    \\bottomrule");
     println!("\\end{{tabular}}");

### 2025-09-13 paper/main.tex
change title

@@
 %\renewcommand{\undertitle}{Technical Report}
-\renewcommand{\shorttitle}{Rust for High-Performance Computing}
+\renewcommand{\shorttitle}{Rust for Scientific Computing}
 
@@
 \hypersetup{
-    pdftitle={Rust for High-Performance Computing},
+    pdftitle={Rust for Scientific Computing},
     pdfsubject={q-bio.NC, q-bio.QM},

### 2025-09-13 paper/main.tex
continue writing introduction

--- 2025-08-30 paper/main.tex
@@
 \hypersetup{
-pdftitle={Rust for High-Performance Computing},
+    pdftitle={Rust for Scientific Computing},
     pdfsubject={q-bio.NC, q-bio.QM},
@@
 
+\printinunitsof{in}\prntlen{\columnwidth}
+
 % keywords can be removed
-\keywords{Rust \and HPC \and High-Performance Computing \and Programming}
+\keywords{Rust \and HPC \and Scientific Computing \and Programming}
 
@@
 \section{Introduction}
-\label{section:introduction}
 
-General Remarks about languages
-\begin{enumerate}
-    \item Programming languages allow us to express machine instructions in human-readable code.
-    \item Depend on use-case
-    \item Translate human-interpretable code into machine instructions
-    \item tradeoff between high-level view (abstraction) and performance (low level)
-    \item C++ has "zero-cost" abstractions
-\end{enumerate}
+% General Remarks about languages
+% \begin{enumerate}
+%     \item Programming languages allow us to express machine instructions in human-readable code.
+%     \item Depend on use-case
+%     \item Translate human-interpretable code into machine instructions
+%     \item tradeoff between high-level view (abstraction) and performance (low level)
+%     \item C++ has "zero-cost" abstractions
+% \end{enumerate}
+
+Programming languages sit at the heart of all things computation.
+They allow users to express higher-level instructions in a human-readable fashion such that this
+code can be translated into machine-level instructions by either compiling or interpreting it.
+Various classes of programming languages have emerged that serve distinct purposes.
+Fundamentally, they aim to descibe problems at various levels of depth which allows humans to
+interpret and reason about the executed instructions and therefore express their desired actions
+more clearly.
 
@@
     \item New fields of study (computational physics, chemistry, biology, sociology, etc.)
-    \item Cite some groundbreaking examples which were possible due to programming in science (maybe weather)
+    \item Cite some groundbreaking examples which were possible due to programming in
+        science (maybe weather)
     \item
 \end{enumerate}
+
+Many languages have been devised to target specific realms which have emerged over the past decades.
+They span across a range of abstraction levels, ranging from low-level languages to high-level ones.
+The low-level operations level mostly deals with instructions on a level close to the actual
+hardware such as memory allocation, pointer arithmetic or by using assembly instructions directly.
+Meanwhile, high-level operations usually combine many lower-level ones to perform a single task.
+An example for a high-level operation would be the addition of two matrices which might allocate the
+required memory, fill it according to the given values and then call the appropriate math function
+to pairwise add their entries.
+Languages which can operate effectively on a lower level often bear the most promise for maximum
+performance which is crucial in any scientific application.
+These languages are mostly compiled using an optimizing compiler which enables many of their
+performance achievements.
+
+This mini-review concerns itself with the Rust programming language and aims to provide an overview
+of the field of scientific computing with the Rust ecosystem.
+Other systems programming languages  such as C, C++ or Fortran have dominated the field over the
+past decades.
+Most notably, C++ has been influential in establishing a model of "zero cost abstractions" where a
+phrasing of the problem from a higher point of view does not involve any additional execution cost
+but is handled by the compiler.
+These and many other principles have allowed for a plethora of
 

### 2025-09-13 paper/main.tex
insert table later

@@
     \centering
-    \includegraphics[width=0.5\textwidth]{figures/crates-io-release-numbers.pdf}
+    \includegraphics[width=0.5\textwidth]{figures/crates-io-release-numbers.pdf}%
+    \includegraphics[width=0.5\textwidth]{figures/crates-io-deps-downloads-scatter.pdf}
     \caption{
@@
         Measure of overall growth of the field.
+        The second figure can be misleading since this will for example not capture the setup where
+        a crate is the dependency of a pypi.org package which is downloaded very often.
+        The colors and size of the markers are also scaled logarithmically with the number of total
+        downloads.
     }
@@
 \section{Ecosystem and Libraries}
+
+\begin{table*}
+    \centering
+    \begin{tabular}{l r r r r r}
+        \toprule
+        Crate Name &Last Update &Version &$N_d$ &$D_\text{week}$ &$D_\text{total}$\\
+        \midrule
+        Lamellar         &2025-03-31 &0.7.1      &42   &20       &     16.9k\\
+        argmin           &2024-02-27 &0.10.0     &18   &12003    &   1227.7k\\
+        arrow            &2025-08-25 &56.1.0     &23   &7532     &  27425.0k\\
+        bevy             &2025-05-30 &0.16.1     &35   &10175    &   2847.7k\\
+        bless            &2025-07-20 &0.1.1      &13   &80       &      0.6k\\
+        burn             &2025-07-18 &0.18.0     &12   &1793     &    319.7k\\
+        c3dio            &2024-04-08 &0.8.0      &2    &24       &     11.8k\\
+        cargo-upgrades   &2025-08-25 &2.2.4      &10   &553      &     64.7k\\
+        cellular\_raza   &2025-08-05 &0.3.0      &10   &157      &     27.4k\\
+        cfsem            &2025-08-14 &2.5.0      &8    &106      &      4.9k\\
+        city2ba          &2020-02-28 &1.1.0      &20   &19       &      3.0k\\
+        cmake            &2025-02-10 &0.1.54     &1    &899615   &  91211.9k\\
+        cudarc           &2025-08-07 &0.17.2     &5    &1184     &    647.0k\\
+        cxxbridge        &2019-12-28 &0.0.0      &0    &14       &      1.3k\\
+        deimos           &2025-06-23 &0.10.0     &13   &40       &      7.1k\\
+        diffsol          &2025-07-16 &0.6.6      &16   &131      &     33.9k\\
+        dioxus           &2025-08-11 &0.6.3      &28   &         &    637.5k\\
+        egobox           &2025-08-22 &0.32.0     &18   &250      &     35.6k\\
+        enzyme           &2021-12-26 &0.4.0      &5    &17       &      5.4k\\
+        extendr-api      &2025-07-12 &0.8.1      &11   &968      &    178.4k\\
+        factrs           &2025-01-13 &0.2.0      &20   &24       &      1.8k\\
+        faer             &2025-04-29 &0.22.6     &46   &1322     &    528.4k\\
+        faucet           &2025-05-31 &0.2.0      &7    &33       &      3.0k\\
+        fem\_2D          &2023-04-20 &0.2.2      &6    &20       &      5.4k\\
+        gcv\_spline      &2024-01-28 &0.2.0      &1    &16       &      2.5k\\
+        \bottomrule
+    \end{tabular}
+    \caption{TODO}
+\end{table*}
 

### 2025-09-15 paper/main.tex
add table package
remove christian from title for now

@@
 \usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{layouts}
 \usepackage{microtype}      % microtypography
 \usepackage{mathrsfs}
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
 \usepackage{graphicx}
@@
     \thanks{
-        \href{https://jonas.pleyer.org}{jonas.pleyer.org},
-        \href{https://cellular-raza.com}{cellular-raza.com}
+        \href{https://jonas.pleyer.org}{jonas.pleyer.org}
     }\\
@@
     University of Freiburg\\
-    \texttt{jonas.pleyer@fdm.uni-freiburg.de} \\
+    \texttt{jonas.pleyer@fdm.uni-freiburg.de}
+    % \\
     %% examples of more authors
-    \And
-    \href{https://orcid.org/0000-0002-6371-4495}{
-        \includegraphics[scale=0.06]{orcid.pdf}
-        \hspace{1mm}Christian Fleck
-    }\\
-    Freiburg Center for Data-Analysis and Modeling\\
-    University of Freiburg
+    % \And
+    % \href{https://orcid.org/0000-0002-6371-4495}{
+    %     \includegraphics[scale=0.06]{orcid.pdf}
+    %     \hspace{1mm}Christian Fleck
+    % }\\
+    % Freiburg Center for Data-Analysis and Modeling\\
+    % University of Freiburg
 }

### 2025-09-15 paper/main.tex
continue writing introduction

@@
 
-Importance of programming in science
-\begin{enumerate}
-    \item New fields of study (computational physics, chemistry, biology, sociology, etc.)
-    \item Cite some groundbreaking examples which were possible due to programming in
-        science (maybe weather)
-    \item
-\end{enumerate}
+Many languages are able to solve many different problems in a general-purpose fashion but some have
+also been developed to target specific realms which have emerged over the past decades.
+Early achievements of the first commercially available programming language FORTRAN are still used
+today in the Modular Ocean Model~\cite{Adcroft2019,Griffies2012}.
+The development of the C programming language in 1972 enabled milestones such as the Unix operating
+system~ \cite{Ritchie1978} which layed the foundation for many major scientific achievements.
+With the advent of C++ in 1985 further contributions were made such as the root project which
+is a "Object Oriented Data-Analysis Framework"~\cite{Brun1997} used by the CERN
+project~\cite{https://doi.org/10.7483/opendata.atlas.l806.5cku}.
+In more recent years  deep learning tools such as pytorch~\cite{Ansel_PyTorch_2_Faster_2024},
+tensorflow~\cite{tensorflow2015-whitepaper} have seen widespread use which are programmed on a
+higher abstraction level using Python.
+The latter enabled the development of AlphaFold which is provides AI-based protein structure
+prediction~\cite{Jumper2021}.
 
-Many languages have been devised to target specific realms which have emerged over the past decades.
-They span across a range of abstraction levels, ranging from low-level languages to high-level ones.
-The low-level operations level mostly deals with instructions on a level close to the actual
-hardware such as memory allocation, pointer arithmetic or by using assembly instructions directly.
+% Importance of programming in science
+% \begin{enumerate}
+%     \item New fields of study (computational physics, chemistry, biology, sociology, etc.)
+%     \item Cite some groundbreaking examples which were possible due to programming in
+%         science (maybe weather)
+%     \item
+% \end{enumerate}
+
+As the preceding exmamples show, even a single programming language can span across multiple
+abstraction levels, ranging from low-levels of operation to high-level ones.
+The former mostly deal with instructions on a level close to the actual hardware such as memory
+allocation, pointer arithmetic or by using assembly instructions directly.
 Meanwhile, high-level operations usually combine many lower-level ones to perform a single task.
@@
 performance which is crucial in any scientific application.
-These languages are mostly compiled using an optimizing compiler which enables many of their
+Of these languages almost all are compiled using an optimizing compiler which enables many of their
 performance achievements.

### 2025-09-15 paper/main.tex
fix typo

@@
 
-As the preceding exmamples show, even a single programming language can span across multiple
+As the preceding examples show, even a single programming language can span across multiple
 abstraction levels, ranging from low-levels of operation to high-level ones.

### 2025-09-24 paper/main.tex
more notes, more formatting, more lazy commit messages

--- 2025-09-15 paper/main.tex
@@
 
+\usepackage{acronym}
 \usepackage[utf8]{inputenc} % allow utf-8 input
@@
 \usepackage{microtype}      % microtypography
+\usepackage{makecell}
 \usepackage{mathrsfs}
 \usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{pagecolor}
 \usepackage{graphicx}
@@
 \usepackage{tikz}
+\usepackage{tabularx}
 \usepackage[backend=biber,style=ieee]{biblatex}
+\usepackage[table]{xcolor}
 
@@
 \usetikzlibrary{trees}
+
+\newacro{raii}[RAII]{Resource Acquisition is Initialization}
+\newacro{ode}[ODE]{Ordinary Differential Equation}
 
@@
 
-\twocolumn
+% \twocolumn
 \pagebreak
 
-\printinunitsof{in}\prntlen{\columnwidth}
-
 % keywords can be removed
-\keywords{Rust \and HPC \and Scientific Computing \and Programming}
+\keywords{Rust \and Scientific Computing \and High-Performance Computing \and Programming}
 
@@
 Various classes of programming languages have emerged that serve distinct purposes.
-Fundamentally, they aim to descibe problems at various levels of depth which allows humans to
+Fundamentally, they aim to descibe problems from different perspectives which allows humans to
 interpret and reason about the executed instructions and therefore express their desired actions
@@
 
-Importance of programming in science
-\begin{enumerate}
-    \item New fields of study (computational physics, chemistry, biology, sociology, etc.)
-    \item Cite some groundbreaking examples which were possible due to programming in
-        science (maybe weather)
-    \item
-\end{enumerate}
+Languages such as C, C++ or Fortran have dominated the field of compiled languages over the past
+decades.
+These languages are able to solve many different problems in a general-purpose fashion.
+Early achievements of the first commercially available programming language Fortran are still used
+today in the Modular Ocean Model~\cite{Adcroft2019,Griffies2012}.
+The development of the C programming language in 1972 enabled milestones such as the Unix operating
+system~ \cite{Ritchie1978} which layed the foundation for many major scientific achievements.
+With the advent of C++ in 1985 further contributions were made such as the root project which
+is an "Object Oriented Data-Analysis Framework"~\cite{Brun1997} and mainly used by
+CERN~\cite{https://doi.org/10.7483/opendata.atlas.l806.5cku}.
+They also build the backbone of all deep learning tools such as
+pytorch~\cite{Ansel_PyTorch_2_Faster_2024} and tensorflow~\cite{tensorflow2015-whitepaper} which
+have seen widespread use and are written on a higher abstraction level using Python.
+Tensorflow builds the foundation of AlphaFold~\cite{Jumper2021} which provides AI-based protein
+structure prediction and Pytorch enabled the development of new cell segmentation
+tools~\cite{Cutler2022}.
 
-Many languages have been devised to target specific realms which have emerged over the past decades.
-They span across a range of abstraction levels, ranging from low-level languages to high-level ones.
-The low-level operations level mostly deals with instructions on a level close to the actual
-hardware such as memory allocation, pointer arithmetic or by using assembly instructions directly.
+% Importance of programming in science
+% \begin{enumerate}
+%     \item New fields of study (computational physics, chemistry, biology, sociology, etc.)
+%     \item Cite some groundbreaking examples which were possible due to programming in
+%         science (maybe weather)
+%     \item
+% \end{enumerate}
+
+A single programming language can span across multiple abstraction levels, ranging from low-levels
+of operation to high-level ones.
+The former mostly deal with instructions close to the actual hardware such as memory allocation,
+pointer arithmetic or by using assembly instructions directly.
 Meanwhile, high-level operations usually combine many lower-level ones to perform a single task.
@@
 Languages which can operate effectively on a lower level often bear the most promise for maximum
-performance which is crucial in any scientific application.
-These languages are mostly compiled using an optimizing compiler which enables many of their
-performance achievements.
+performance which is crucial for most scientific applications.
+% Of these languages almost all are compiled using an optimizing compiler which enables many of their
+% performance achievements.
 
-This mini-review concerns itself with the Rust programming language and aims to provide an overview
-of the field of scientific computing with the Rust ecosystem.
-Other systems programming languages  such as C, C++ or Fortran have dominated the field over the
-past decades.
-Most notably, C++ has been influential in establishing a model of "zero cost abstractions" where a
-phrasing of the problem from a higher point of view does not involve any additional execution cost
-but is handled by the compiler.
-These and many other principles have allowed for a plethora of
+This review concerns itself with the Rust programming language and aims to provide an overview
+of the field of scientific computing within the Rust ecosystem.
+Rust supports multiple programming paradigms and offers built-in memory safety by employing an
+ownership model that enforces the validity of all references.
+As such it is able to cover a wide range of abstraction levels, comparable with C++ and offers
+comparable performance \textbf{CITATION}.
+It takes hard-learned lessons and offers solutions for many existing problems that other languages
+face.
+It has been the "most loved language" for the past 7 years according to data from the developer
+survey of \url{stackoverflow.com}.
+This data is shown in figure~\ref{fig:stackoverflow-popular-languages} where we compare it to other
+popular programming languages is compared against other languages.
+It also ranks high among the "most desired languages" which is displayed in
+figure~\ref{fig:stackoverflow-desired-languages}.
 
-Judging C/C++
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{figures/stackoverflow-loved-language.pdf}%
+    \includegraphics[width=0.5\textwidth]{figures/stackoverflow-desired-language.pdf}
+    \caption{
+        Popularity of the Rust programming language.
+    }
+    \label{fig:stackoverflow-desired-languages}
+\end{figure}
+
+\subsection{Comparison of Rust with C and C++}
+Within the realm of scientific computing, C++ and C can be seen as the major competitors of Rust.
+As previously discussed, they have been instrumental in the achievement of many scientific
+milestones.
+They are widely applied and developers are generally readily available to tackle new projects.
+Thus, their importance to the foundational advances across all areas of computation should not be
+understated.
+However, as with any maturing technology, many issues have been identified which are intrinsic to
+the design of these languages.
+The C programming language manually manages the memory of the program which provides flexibility but
+leaves the task of correctly freeing resources to the user which can quickly become difficult to
+track.
+This ambiguity has led to many bugs and vulnerabilities.
+The chromium project has published data that indicates $\approx 70\%$ of all high-severity bugs to
+be related to memory unsafety~\cite{ChromiumMemorySafety}.
+These issues are important for all practical applications, especially at larger scales with
+multiple workstations, clusters or supercomputers where program crashes can lead to lost data and
+waisted resources.
+
+Introduce \ac{raii}
+
+\begin{table}
+    \centering
+    \begin{tabularx}{\textwidth}{X X X X}
+        % \toprule
+        Category & C & C++ & Rust\\
+        \midrule
+        Memory Safety & Dangling pointers, buffer overflows, manual memory management & Includes C,
+        mitigated by \ac{raii} \& smart pointers & Ownership model\\
+        \midrule
+        % Undefined Behavior & 
+        Error Handling & Return error codes & Return error codes \& Exceptions &
+        Errors are return types, \texttt{Result} and \texttt{Option} types enforce handling\\
+        \midrule
+        \makecell[l]{Concurrency \&\\ Parallelism} & Manual (\texttt{pthreads})
+        & Manual, risk data races &
+        Thread-safety through ownership model, enforced at compile-time\\
+        \midrule
+        Ecosystem & Extensive, fragmented & Vast, fragmented (CMake, Conan, vcpkg, ..) & Unified
+        package manager, registry, build system\\
+        % \bottomrule
+    \end{tabularx}
+\end{table}
+
+Since C++ provides backwards compatiblity
+
+C++ has been influential in establishing a model of "zero cost abstractions" where a phrasing of the
+problem from a higher point of view does not involve any additional execution cost but is handled by
+the compiler.
+
+\begin{enumerate}
+    \item Memory Safety Issues
+    \item Mutable by Default
+    \item Macros
+    \item Ecosystem Fragmentation
+    \item Concurrency and Parallelism
+    \item Error Handling
+\end{enumerate}
+
+C++
+\begin{enumerate}
+    \item Legacy and Backward Compatibility
+    \item Templates vs. Generics
+    \item Complexity of the Standard
+    \item Multiple Inheritance Complexity
+    \item Name Lookup and Scope Complexity
+    \item Undefined Behavior is Easier to Trigger
+\end{enumerate}
+
 \begin{enumerate}
@@
 
+\subsection{Current Advancements of Rust}
 Show that many companies are starting to adopt Rust
@@
 
-\begin{figure*}
+\begin{figure}
     \centering
@@
         Usage of programming languages according to (A) stackoverflow.com and (B) pypi.org.
+        % For discussion: overlap between C and C++ is about 406 projects
+        % Other overlaps are negligible
     }
-\end{figure*}
-\begin{figure*}
-    \centering
-    \includegraphics[width=0.5\textwidth]{figures/stackoverflow-loved-language.pdf}%
-    \includegraphics[width=0.5\textwidth]{figures/stackoverflow-desired-language.pdf}
-    \caption{
-        Popularity of the Rust programming language.
-    }
-\end{figure*}
-\begin{figure*}
+    \label{fig:stackoverflow-popular-languages}
+\end{figure}
+\begin{figure}
     \centering
@@
     }
-\end{figure*}
+\end{figure}
 
@@
 
-\begin{table*}
+\begin{table}
     \centering
+    \rowcolors{2}{white}{lightgray!30}
     \begin{tabular}{l r r r r r}
@@
     \caption{TODO}
-\end{table*}
+\end{table}
 
@@
 
-\onecolumn
+% \onecolumn
 \printbibliography

### 2025-09-24 figures/pypi-org-used-languages.py
correct data

@@ -29,12 +29,12 @@ def prepare_fig_ax():
 
 
 data = [
-    ("Go", 20 * 3, "#00acd7"),
-    ("Java", 20 * 10, "#c12f30"),
-    ("Fortran", 20 * 13, "#744e97"),
-    ("C", 20 * 101, "#659bd3"),
-    ("C++", 20 * 119, "#00599d"),
-    ("Rust", 20 * 158, "#CE412B"),
+    ("Go", 46, "#00acd7"),
+    ("Java", 186, "#c12f30"),
+    ("Fortran", 248, "#744e97"),
+    ("C", 2013, "#659bd3"),
+    ("C++", 2403, "#00599d"),
+    ("Rust", 3244, "#CE412B"),
 ]
 
 fig, ax = prepare_fig_ax()
